### 2018-08-15 Assignment_1/report.tex
Add data tracking capability

\documentclass[11pt,a4paper]{article}

\usepackage{fullpage}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

\begin{document}

\title{COSC3500 \\ 2D Orbital Simulation Report}
\author{Maxwell Bo (43926871)}
\date{August 17, 2018}
\maketitle

% Description of problem - readable and covers all salient points - /2
\subsection*{Description}

% Description of implementation - is coherent and has sufficient depth and detail - /5
\subsection*{Implementation}

% Why do you believe it is correct - /3
\subsection*{Correctness}

% Scaling or other performance discussion - /2
\subsection*{Performance \& Scaling}

The addition of the 
\texttt{-march=native} compiler flag, which enables the use of all CPU specific instructions, provided no recognizable improvement in running time.

The use of the GCC and Clang Profile-Guided Optimisation features provided no recognizable improvement in running time.

Distressingly, \texttt{-O0}, \texttt{-O1}, \texttt{-O2}, \texttt{-O3} showed no recognizable improvement in running time. \texttt{-Ofast} led to an -4\%-ish performance regression.

\includegraphics[width=\textwidth]{profile}

% The \texttt{exert_force_unidirectionally} method was an incredibly costly method, with an exclusive cost of 26.33\% of total running time.

Two performance fixes were divised.

\begin{lstlisting}
void Body::exert_force_unidirectionally(const Body& there) {
    double m1 = m;
    double m2 = there.m;
    double r = distance(x, y, there.x, there.y);
    double r2 = r * r;

    double F = (G * m1 * m2) / r2;

    double delta_x = there.x - x;
    double delta_y = there.y - y;

    // turn the displacement vector between our two points into a force vector
    // of the desired magnitude
    double scale_factor = F / r;

    Fx += delta_x * scale_factor;
    Fy += delta_y * scale_factor;
}
\end{lstlisting}

Here, we are recalculating

gcc -g -lstdc++ -Wall -pedantic -Wextra -std=c++11 -lm -O3 -march=native Body.o
QuadTree.o main.o utils.o -o nbody
Barnes-Hut enabled: false
Leapfrog enabled: true
Total CPU time was 13.938825
12000001 simulation steps computed

\end{document}

### 2018-08-15 Assignment_1/src/main.cpp

@@ -210,6 +210,7 @@ int main(int argc, char **argv) {
     const std::string &input_filename = argv[4];
 
     std::ifstream input_fh(input_filename);
+    FILE *log_fh = fopen("nbody.log", "a");
     std::vector<Body> bodies = {};
 
     parse_input_file(input_fh, bodies);
@@ -291,8 +292,20 @@ int main(int argc, char **argv) {
         }
     }
 
-    fprintf(stderr, "Total CPU time was %lf\n", cpu_time() - start);
-    fprintf(stderr, "%d simulation steps computed\n", step);
+    const double cpu_time_elapsed = cpu_time() - start;
+
+    fprintf(
+        log_fh,
+        ",\n{ 'numTimeSteps': %d, 'inputFile': %s, 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
+        num_time_steps,
+        input_filename.c_str(),
+        static_cast<int>(bodies.size()), // thank-you Joel
+        cpu_time_elapsed,
+        ENABLE_LEAPFROG ? "true" : "false",
+        ENABLE_BARNES_HUT ? "true" : "false"
+    );
+
+    fprintf(stderr, "Total CPU time was %lf\n", cpu_time_elapsed);
     
     return 0;
 }

### 2018-08-15 Assignment_1/src/main.cpp
Add a logging framework

@@ -16,7 +16,8 @@ const unsigned int LEAP = 0;
 const unsigned int FROG = 1;
 
 const bool ENABLE_BARNES_HUT = false;
-const bool ENABLE_LEAPFROG = true;
+const bool ENABLE_LEAPFROG = false;
+const bool ENABLE_LOGGING = true;
 
 double cpu_time(void) {
     return (double)clock() / (double)CLOCKS_PER_SEC;
@@ -294,18 +295,21 @@ int main(int argc, char **argv) {
 
     const double cpu_time_elapsed = cpu_time() - start;
 
-    fprintf(
-        log_fh,
-        ",\n{ 'numTimeSteps': %d, 'inputFile': %s, 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
-        num_time_steps,
-        input_filename.c_str(),
-        static_cast<int>(bodies.size()), // thank-you Joel
-        cpu_time_elapsed,
-        ENABLE_LEAPFROG ? "true" : "false",
-        ENABLE_BARNES_HUT ? "true" : "false"
-    );
+    if (ENABLE_LOGGING) {
+        fprintf(
+            log_fh,
+            ",\n{ 'numTimeSteps': %d, 'inputFile': %s, 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
+            num_time_steps,
+            input_filename.c_str(),
+            static_cast<int>(bodies.size()), // thank-you Joel
+            cpu_time_elapsed,
+            ENABLE_LEAPFROG ? "true" : "false",
+            ENABLE_BARNES_HUT ? "true" : "false"
+        );
+    }
 
     fprintf(stderr, "Total CPU time was %lf\n", cpu_time_elapsed);
+    fclose(log_fh);
     
     return 0;
 }

### 2018-08-15 Assignment_1/run.sh
Update log output

@@ -1,1 +1,1 @@
-make && ./nbody 6000000 0.016666666666666666 10 $1 > outputs/out.data
+make && ./nbody 600000 0.016666666666666666 10 $1 > outputs/out.data

### 2018-08-15 Assignment_1/src/main.cpp

@@ -16,7 +16,7 @@ const unsigned int LEAP = 0;
 const unsigned int FROG = 1;
 
 const bool ENABLE_BARNES_HUT = false;
-const bool ENABLE_LEAPFROG = false;
+const bool ENABLE_LEAPFROG = true;
 const bool ENABLE_LOGGING = true;
 
 double cpu_time(void) {
@@ -298,7 +298,7 @@ int main(int argc, char **argv) {
     if (ENABLE_LOGGING) {
         fprintf(
             log_fh,
-            ",\n{ 'numTimeSteps': %d, 'inputFile': %s, 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
+            ",\n{ 'numTimeSteps': %d, 'inputFile': '%s', 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
             num_time_steps,
             input_filename.c_str(),
             static_cast<int>(bodies.size()), // thank-you Joel

### 2018-08-16 Assignment_1/src/main.cpp
Use move semantics for packed vector

@@ -125,10 +125,8 @@ x1 y1 vx1 vy1
 ..
 xN yN vxN vyN
 */
-void parse_input_file(
-    std::ifstream& input_fh,
-    std::vector<Body>& bodies
-) {
+std::vector<Body> parse_input_file(std::ifstream& input_fh) {
+    std::vector<Body> bodies = {};
     std::vector<double> masses = {};
     std::string line;
 
@@ -184,12 +182,16 @@ void parse_input_file(
         input_fh.close();
     }
 
+    // we're going to throw bodies_n away here and let the Body vector
+    // be the source of truth for how many we have
     assert(bodies_n == bodies.size());
     assert(bodies_n == masses.size());
 
     for (size_t i = 0; i < bodies.size(); i++) {
         bodies[i].m = masses[i];
     }
+
+    return bodies;
 }
 
 int main(int argc, char **argv) {
@@ -212,9 +214,8 @@ int main(int argc, char **argv) {
 
     std::ifstream input_fh(input_filename);
     FILE *log_fh = fopen("nbody.log", "a");
-    std::vector<Body> bodies = {};
 
-    parse_input_file(input_fh, bodies);
+    std::vector<Body> bodies = parse_input_file(input_fh);
 
     double t = 0; // XXX: optimization - double source of truth, update both
     // TODO: can we have a function that inlines and updates both of these?

### 2018-08-16 Assignment_1/viz.py
Update viz

@@ -6,6 +6,7 @@ import math
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.animation as animation
+import matplotlib.gridspec as gridspec
 
 EDGE = 125
 SHOULD_EXPORT = False
@@ -39,6 +40,9 @@ def main():
         exit(1)
 
     with open(sys.argv[1], "r") as f:
+        ###########
+        # PARSING #
+        ###########
         contents = f.read()
 
         bodies_n, timestep_n, interval, delta_t = contents.split('\n')[0].split()
@@ -51,40 +55,25 @@ def main():
         timestep_block = "\n".join(contents.split('\n')[bodies_n + 1:])
         timesteps = [ parse_timestep(i) for i in timestep_block.strip().split("\n\n") ]
 
-        # import numpy as np
-        # import matplotlib.pyplot as plt 
-        # from matplotlib import gridspec
+        ##########
+        # LAYOUT #
+        ##########
+        fig = plt.figure(figsize=(8, 10))
+        gs = gridspec.GridSpec(2, 1, height_ratios=[3, 1]) 
 
-        # # generate some data
-        # x = np.arange(0, 10, 0.2)
-        # y = np.sin(x)
-
-        # # plot it
-        # fig = plt.figure(figsize=(8, 6)) 
-        # gs = gridspec.GridSpec(1, 2, width_ratios=[3, 1]) 
-        # ax0 = plt.subplot(gs[0])
-        # ax0.plot(x, y)
-        # ax1 = plt.subplot(gs[1])
-        # ax1.plot(y, x)
-
-        # plt.tight_layout()
-        # plt.savefig('grid_figure.pdf')
-
-        fig = plt.figure(figsize=(8, 12))
-
-        ax1 = fig.add_subplot(211)
-        ax2 = fig.add_subplot(212)
-
+        ax1 = plt.subplot(gs[0])
         ax1.set_xlim(-EDGE, EDGE)
         ax1.set_ylim(-EDGE, EDGE)
 
-        # ax1.figure.set_size_inches(8, 8, forward=True)
+        ax2 = plt.subplot(gs[1])
 
+        plt.tight_layout()
+
+        ############
+        # PLOTTING #
+        ############
         sizes = [ (math.log(mass / 1e14) + 1) * 10 for mass in masses ]
         colors = np.random.rand(bodies_n)
-
-        cursor = ax2.axvline(x = 0, animated=True)
-
         particles = ax1.scatter(
             [particle.x for particle in timesteps[0].bodies], 
[…]

### 2018-08-16 Assignment_1/src/main.cpp
Use shared_ptrs for Barnes-hut

@@ -15,7 +15,7 @@
 const unsigned int LEAP = 0;
 const unsigned int FROG = 1;
 
-const bool ENABLE_BARNES_HUT = false;
+const bool ENABLE_BARNES_HUT = true;
 const bool ENABLE_LEAPFROG = true;
 const bool ENABLE_LOGGING = true;
 

### 2018-08-16 Assignment_1/viz.py
Undo accidental alterations

@@ -58,7 +58,7 @@ def main():
         ##########
         # LAYOUT #
         ##########
-        fig = plt.figure(figsize=(8, 10))
+        fig = plt.figure(figsize=(6, 8))
         gs = gridspec.GridSpec(2, 1, height_ratios=[3, 1]) 
 
         ax1 = plt.subplot(gs[0])

### 2018-08-16 Assignment_1/viz.py
Drop marker size

@@ -84,7 +84,7 @@ def main():
 
         cursor = ax2.axvline(x = 0, animated=True)
         frames = np.arange(0, len(timesteps))
-        ax2.plot(frames, [timestep.total_energy for timestep in timesteps], "bo")
+        ax2.plot(frames, [timestep.total_energy for timestep in timesteps], "bo", markersize=2)
 
         def init():
             return particles, cursor

### 2018-08-16 Assignment_1/src/main.cpp
Update ./run.sh

@@ -17,7 +17,7 @@ const unsigned int FROG = 1;
 
 const bool ENABLE_BARNES_HUT = true;
 const bool ENABLE_LEAPFROG = true;
-const bool ENABLE_LOGGING = true;
+const bool ENABLE_LOGGING = false;
 
 double cpu_time(void) {
     return (double)clock() / (double)CLOCKS_PER_SEC;

### 2018-08-16 Assignment_1/src/QuadTree.cpp
Readd QuadTree shared_ptr impl

@@ -13,18 +13,19 @@ const double THETA = 0.5;
 QuadTree::QuadTree() {
 }
 
-QuadTree::QuadTree(double x, double y, double radius):
+QuadTree::QuadTree(double x, double y, double radius): 
     x(x),
     y(y),
     radius(radius),
     mx(0),
     my(0),
     m(1),
-    ne(nullptr),
-    nw(nullptr),
-    se(nullptr),
-    sw(nullptr),
-    occupant(nullptr) {
+    ne(std::shared_ptr<QuadTree>()),
+    nw(std::shared_ptr<QuadTree>()),
+    se(std::shared_ptr<QuadTree>()),
+    sw(std::shared_ptr<QuadTree>()),
+    occupant(nullptr)
+{
 }
 
 bool QuadTree::within_bounds(const Body& body) const {
@@ -44,10 +45,10 @@ void QuadTree::subdivide() {
     const double r = radius / 2;
 
     // (x, y, radius), love me positional calling semantics
-    ne = new QuadTree(x + r, y + r, r);
-    nw = new QuadTree(x - r, y + r, r);
-    se = new QuadTree(x + r, y - r, r);
-    sw = new QuadTree(x - r, y - r, r);
+    ne = std::make_shared<QuadTree>(x + r, y + r, r);
+    nw = std::make_shared<QuadTree>(x - r, y + r, r);
+    se = std::make_shared<QuadTree>(x + r, y - r, r);
+    sw = std::make_shared<QuadTree>(x - r, y - r, r);
 }
 
 /*  To calculate the net force acting on body b, use the following recursive procedure, 
@@ -132,7 +133,6 @@ void QuadTree::calculate_force(Body& body) {
         there may be several subdivisions during a single insertion.
         Finally, update the center-of-mass and total mass of x.
 */
-
 bool QuadTree::insert_all(std::vector<Body>& bodies) {
     for (auto& body : bodies) {
         const bool did_insert = insert(body);

### 2018-08-16 Assignment_1/src/QuadTree.hpp

@@ -14,10 +14,10 @@ class QuadTree {
         double mx;
         double my;
         double m;
-        QuadTree* ne;
-        QuadTree* nw;
-        QuadTree* se;
-        QuadTree* sw;
+        std::shared_ptr<QuadTree> ne;
+        std::shared_ptr<QuadTree> nw;
+        std::shared_ptr<QuadTree> se;
+        std::shared_ptr<QuadTree> sw;
         Body *occupant;
         // Methods
         bool insert(Body& body);

### 2018-08-16 Assignment_1/run.sh
Run some tests

@@ -1,5 +1,6 @@
 hash=`git rev-parse HEAD`
 name="$hash"
 
-make && ./nbody 600000 0.016666666666666666 10 $1 > "outputs/$name.data" && python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
+make && ./nbody 6000000 0.016666666666666666 10 $1 > "outputs/$name.data" 
+# python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-16 Assignment_1/src/main.cpp

@@ -15,9 +15,9 @@
 const unsigned int LEAP = 0;
 const unsigned int FROG = 1;
 
-const bool ENABLE_BARNES_HUT = true;
+const bool ENABLE_BARNES_HUT = false;
 const bool ENABLE_LEAPFROG = true;
-const bool ENABLE_LOGGING = false;
+const bool ENABLE_LOGGING = true;
 
 double cpu_time(void) {
     return (double)clock() / (double)CLOCKS_PER_SEC;

### 2018-08-16 Assignment_1/src/Body.cpp
Precompute Gm

@@ -5,10 +5,9 @@
 #include "Body.hpp"
 #include "utils.hpp"
 
-const double G = 6.674e-11; // N(m/kg) ^ 2
-
 Body::Body():
     m(1),
+    Gm(G * m),
     x(0),
     y(0),
     vx(0),
@@ -83,11 +82,11 @@ double Body::kinetic_energy() const {
 double Body::gravitational_potential_energy(const Body& there) const {
     const double R = distance(x, y, there.x, there.y); // final distance, aka, to edge
 
-    return (-G * m * there.m) / R;
+    return (-Gm * there.m) / R;
 }
 
 void Body::exert_force_unidirectionally(const Body& there) {
-    const double m1 = m;
+    // const double m1 = m; We have a precomputed Gm that we're going to use
     const double m2 = there.m;
 
     // Rather than calling `distance`, we re-use Δx and Δy
@@ -101,7 +100,7 @@ void Body::exert_force_unidirectionally(const Body& there) {
     const double r = hypot(delta_x, delta_y);
     const double r2 = pow(r, 2);
 
-    const double F = (G * m1 * m2) / r2;
+    const double F = (Gm * m2) / r2;
 
     // turn the displacement vector between our two points into a force vector
     // of the desired magnitude
@@ -113,7 +112,6 @@ void Body::exert_force_unidirectionally(const Body& there) {
 
 void Body::exert_force_bidirectionally(Body& there) {
     // See exert_force_unidirectionally for inline commentary
-    const double m1 = m;
     const double m2 = there.m;
 
     const double delta_x = there.x - x;
@@ -122,7 +120,7 @@ void Body::exert_force_bidirectionally(Body& there) {
     const double r = hypot(delta_x, delta_y);
     const double r2 = pow(r, 2);
 
-    const double F = (G * m1 * m2) / r2;
+    const double F = (Gm * m2) / r2;
 
     const double scale_factor = F / r;
 

### 2018-08-16 Assignment_1/src/Body.hpp

@@ -1,9 +1,12 @@
 #ifndef _Body_h
 #define _Body_h
 
+const double G = 6.674e-11; // N(m/kg) ^ 2
+
 class Body {
     public:
         double m;
+        double Gm;
         // Position
         double x;
         double y;

### 2018-08-16 Assignment_1/src/main.cpp

@@ -189,6 +189,7 @@ std::vector<Body> parse_input_file(std::ifstream& input_fh) {
 
     for (size_t i = 0; i < bodies.size(); i++) {
         bodies[i].m = masses[i];
+        bodies[i].Gm = G * masses[i];
     }
 
     return bodies;
@@ -299,7 +300,7 @@ int main(int argc, char **argv) {
     if (ENABLE_LOGGING) {
         fprintf(
             log_fh,
-            ",\n{ 'numTimeSteps': %d, 'inputFile': '%s', 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s }",
+            ",\n{ 'numTimeSteps': %d, 'inputFile': '%s', 'numBodies': %d, 'time': %lf, 'leapfrog': %s, 'barnesHut': %s, 'precomputedGm': true }",
             num_time_steps,
             input_filename.c_str(),
             static_cast<int>(bodies.size()), // thank-you Joel

### 2018-08-16 Assignment_1/viz.py
Update labels

@@ -62,10 +62,13 @@ def main():
         gs = gridspec.GridSpec(2, 1, height_ratios=[3, 1]) 
 
         ax1 = plt.subplot(gs[0])
+        ax1.set_title(sys.argv[1])
         ax1.set_xlim(-EDGE, EDGE)
         ax1.set_ylim(-EDGE, EDGE)
+        ax1.set(xlabel="x (m)", ylabel="y (m)")
 
         ax2 = plt.subplot(gs[1])
+        ax2.set(xlabel="t (s)", ylabel="Energy (J)")
 
         plt.tight_layout()
 
@@ -83,14 +86,17 @@ def main():
         )
 
         cursor = ax2.axvline(x = 0, animated=True)
+        fps = round(1 / interval)
         frames = np.arange(0, len(timesteps))
-        ax2.plot(frames, [timestep.total_energy for timestep in timesteps], "bo", markersize=2)
+        time = np.true_divide(frames, fps)
+
+        ax2.plot(time, [timestep.total_energy for timestep in timesteps], "bo", markersize=2)
 
         def init():
             return particles, cursor
 
-        def animate(t):
-            timestep = timesteps[t]
+        def animate(step):
+            timestep = timesteps[step]
 
             xs = [particle.x for particle in timestep.bodies]
             ys = [particle.y for particle in timestep.bodies]
@@ -99,7 +105,7 @@ def main():
 
             particles.set_offsets(xys)
 
-            cursor.set_xdata(t)
+            cursor.set_xdata(step / fps)
 
             return particles, cursor
 
@@ -107,7 +113,6 @@ def main():
                                     interval=interval * 1000, blit=True)
 
         if len(sys.argv) == 1 + 2:
-            fps = round(1 / interval)
 
             print("Exporting at", fps, "FPS")
             ani.save(sys.argv[2], writer=animation.FFMpegWriter(

### 2018-08-16 Assignment_1/viz.py
Update viz labels

@@ -68,7 +68,7 @@ def main():
         ax1.set(xlabel="x (m)", ylabel="y (m)")
 
         ax2 = plt.subplot(gs[1])
-        ax2.set(xlabel="t (s)", ylabel="Energy (J)")
+        ax2.set(xlabel="t (s)", ylabel="E (J)")
 
         plt.tight_layout()
 

### 2018-08-16 Assignment_1/src/main.cpp
Fix time stepping error

@@ -286,7 +286,7 @@ int main(int argc, char **argv) {
                 body.euler_integrate(timestep);
             }
 
-            t += halfstep;
+            t += timestep;
             step++;
         }
 

### 2018-08-16 Assignment_1/run.sh
Fix parsing regime

@@ -1,6 +1,6 @@
-hash=`git rev-parse HEAD`
-name="$hash"
+hash=`git rev-parse --short HEAD`
+name="$1-$hash"
 
-make && ./nbody 6000000 0.016666666666666666 10 $1 > "outputs/$name.data" 
-# python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
+make && ./nbody 6000000 0.016666666666666666 10 "inputs/$1" > "outputs/$name.data" 
+python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-16 Assignment_1/report.tex
Update report

@@
 \usepackage[utf8]{inputenc}
+\usepackage{amsmath}
 
@@
 \subsection*{Description}
+The task was to create a stock-standard, 2-dimensional gravitational $n$-body simulator. 
+All bodies were to be assumed to be point masses. The simulation was to be accurate, maintaining a constant total energy, and exhibiting phenomena such as apsidal precession. 
+The simulation was to accept arguments specifying the granularity of the simulation (number of time steps, number of instances of data export), and a file specifiying masses and their initial positions and velocities.
+The simulation was to produce its output as fast as possible, with minimal slowdown with an increasing $n$ number of bodies.
+
+The simulator did not need to handle collisions between bodies.
+The MS1 simulator was to be free of OpenMP multiprocessing, which would be implemented in preparation for the MS2 submission.
 
@@
 
+At a high-level, the initial naive simulator:
+
+\begin{enumerate}
+    \item Parsed input parameters and files
+    \item Constructed \texttt{Body} class instances, representing each point mass
+    \item Packed the \texttt{Body}s into a \texttt{std::vector<Body>}, to maximize cache locality
+    \item Calculated forces between all pairwise combinations of $n$-bodies ($O(n^2)$)
+    \item Performed Euler's method to derive new velocities and positions
+    \item Output all necessary data
+    \item \texttt{GOTO 4}
+\end{enumerate}
+
+By using a Quadtree, (`a tree datastructure in which each internal node has exactly four children') and the \textit{Barnes-Hut} algorithm\cite{barnes86a}, the total cost of force calculation could be reduced to $O(n \log n)$, by grouping close-together bodies and approximating forces between the singular grouped pseudo-body, and distant bodies.
+New Quadtrees were constructed on each seperate simulation step.
+
+Dehen and Read note that the Euler method `performs very poorly in practice', further noting that `errors are proportional to $\Delta t^2$'.
+They contrast it with the second-order leapfrog symplectic integrator, which is `heavily used in collisionless N-body applications'.\cite{dehen11a}
+
+Leapfrog can be expressed many in forms\cite{Skeel1993} including a synchronised form:
+
+\begin{align*}
+x_{i}&=x_{i-1}+v_{i-1/2}\,\Delta t\\
+a_{i}&=F(x_{i})\\
+v_{i+1/2}&=v_{i-1/2}+a_{i}\,\Delta t\\
+\end{align*}
+
+which only requires a single acceleration calculation per every two half timesteps (the timestep $\Delta t$ must be constant to maintain stability), and a `kick-drift-kick' form
+
+\begin{align*}
+v_{i+1/2}&=v_{i}+a_{i}{\frac {\Delta t}{2}}\\
+x_{i+1}&=x_{i}+v_{i+1/2}\Delta t\\
+v_{i+1}&=v_{i+1/2}+a_{i+1}{\frac {\Delta t}{2}}
+\end{align*}
+
+that is stable with variable timstepping, but incurs an additional acceleration calculation per every two half timesteps.
+
+The synchronised form was implemented, but attempts to implement the kick-drift-kick form, and variable timestepping, were left unfinished.
+
+Thus, the final implementation:
+
+\begin{enumerate}
+    \item Parsed input parameters and files
+    \item Constructed \texttt{Body} class instances, representing each point mass
+    \item Packed the \texttt{Body}s into a \texttt{std::vector<Body>}, to maximize cache locality
+    \item Inserted all \texttt{Body}s into a fresh \texttt{QuadTree} on full timesteps, traversing the \texttt{QuadTree} with every \texttt{Body} to calculate forces ($O(n \log n)$)
+    \item Performed the appropriate Leapfrog step to derive new velocities \textit{or} positions
+    \item Output all necessary data
+    \item \texttt{GOTO 4}
+\end{enumerate}
+
 % Why do you believe it is correct - /3
 \subsection*{Correctness}
+
+I personally believe that the simulation is \textit{mostly} right.
+By visualising the results with \texttt{matplotlib}, we see something that resembles an $n$-body simulator\footnote{and that's good enough for me}.
+Distressingly, the total energy of the system is not constant throughout the system.
+When bodies are in close proxmity, anomalous energies that are inconsistent with the energy curve are observed.
+Furthermore, simulations with higher numbers of bodies produce random low energy outliers, but have a smoother total energy curve.
+There seems to be no signfiicant difference between Euler method and Leapfrog, with respect to energy anomalies.
+
+\begin{figure}[b]
+\caption{Observed energy anomaly while bodies in close proxmity - Leapfrog}
+\centering
+\includegraphics[width=0.5\textwidth]{energy_anomaly}
+\end{figure}
+
+\begin{figure}[b]
+\caption{Observed energy anomaly while bodies in close proxmity - Euler method}
+\centering
+\includegraphics[width=0.5\textwidth]{euler_energy_anomaly}
+\end{figure}
+
+\begin{figure}[b]
+\caption{Observed energy anomaly while bodies in close proxmity - Euler method}
+\centering
+\includegraphics[width=0.5\textwidth]{euler_relaxed_energy_anomaly}
+\end{figure}
+
+\begin{figure}[b]
+\caption{Observed energy anomaly - random outlier - Leapfrog}
+\centering
+\includegraphics[width=0.5\textwidth]{relaxed_energy_anomaly}
+\end{figure}
 
@@
 
+Henceforth, the use of the `recognizable' refers to eyeballing the output data, and making no significant effort to investigate the data more rigorously.
+
 The addition of the 
-\texttt{-march=native} compiler flag, which enables the use of all CPU specific instructions, provided no recognizable improvement in running time.
+\texttt{-march=native} compiler flag, which enables the use of all CPU specific instructions, provided no recognizable improvement in running time, but was left enabled in the instance that it improved performance on \texttt{goliath}.
 
-The use of the GCC and Clang Profile-Guided Optimisation features provided no recognizable improvement in running time.
+The use of both th GCC and Clang Profile-Guided Optimisation features provided no recognizable improvement in running time.
 
 Distressingly, \texttt{-O0}, \texttt{-O1}, \texttt{-O2}, \texttt{-O3} showed no recognizable improvement in running time. \texttt{-Ofast} led to an -4\%-ish performance regression.
+
+By profiling with \texttt{callgrind}, we saw that execution was dominated by only one user-defined method.
 
@@
 
-\begin{lstlisting}
-void Body::exert_force_unidirectionally(const Body& there) {
-    double m1 = m;
-    double m2 = there.m;
-    double r = distance(x, y, there.x, there.y);
-    double r2 = r * r;
-
-    double F = (G * m1 * m2) / r2;
-
-    double delta_x = there.x - x;
-    double delta_y = there.y - y;
-
-    // turn the displacement vector between our two points into a force vector
-    // of the desired magnitude
-    double scale_factor = F / r;
-
-    Fx += delta_x * scale_factor;
-    Fy += delta_y * scale_factor;
-}
-\end{lstlisting}
+% \begin{lstlisting}
+% \end{lstlisting}
 
@@
 
+\medskip
+
+\bibliographystyle{ieeetr}
+\bibliography{bib}
+
 \end{document}

### 2018-08-16 Assignment_1/src/main.cpp

@@ -17,7 +17,7 @@ const unsigned int FROG = 1;
 
 const bool ENABLE_BARNES_HUT = false;
 const bool ENABLE_LEAPFROG = true;
-const bool ENABLE_LOGGING = true;
+const bool ENABLE_LOGGING = false;
 
 double cpu_time(void) {
     return (double)clock() / (double)CLOCKS_PER_SEC;
@@ -126,6 +126,7 @@ x1 y1 vx1 vy1
 xN yN vxN vyN
 */
 std::vector<Body> parse_input_file(std::ifstream& input_fh) {
+    // not a Body&; we want to maximise cache locality
     std::vector<Body> bodies = {};
     std::vector<double> masses = {};
     std::string line;

### 2018-08-16 Assignment_1/viz.py

@@ -39,7 +39,9 @@ def main():
         print("input [output]")
         exit(1)
 
-    with open(sys.argv[1], "r") as f:
+    in_filename = sys.argv[1]
+
+    with open(in_filename, "r") as f:
         ###########
         # PARSING #
         ###########
@@ -47,7 +49,7 @@ def main():
 
         bodies_n, timestep_n, interval, delta_t = contents.split('\n')[0].split()
         bodies_n, timestep_n, interval, delta_t = int(bodies_n), int(timestep_n), float(interval), float(delta_t)
-        print({"numBodies": bodies_n, "numTimeSteps": timestep_n, "outputInterval": interval, "deltaT": delta_t})
+        print({"input": in_filename, "numBodies": bodies_n, "numTimeSteps": timestep_n, "outputInterval": interval, "deltaT": delta_t})
 
         masses = [ float(i) for i in contents.split('\n')[1:bodies_n + 1] ]
         assert(len(masses) == bodies_n)
@@ -114,8 +116,10 @@ def main():
 
         if len(sys.argv) == 1 + 2:
 
+            out_filename = sys.argv[2]
+
             print("Exporting at", fps, "FPS")
-            ani.save(sys.argv[2], writer=animation.FFMpegWriter(
+            ani.save(out_filename, writer=animation.FFMpegWriter(
                 fps=fps, 
                 metadata={
                     "artist": "Max Bo",

### 2018-08-16 Assignment_1/report.tex
Update report

@@
 
-By profiling with \texttt{callgrind}, we saw that execution was dominated by only one user-defined method.
+By profiling with \texttt{callgrind}, we saw that execution was dominated by only one incredibly costly user-defined method, with an exclusive cost of 26.33\% of total running time.
 
@@
 
-% The \texttt{exert_force_unidirectionally} method was an incredibly costly method, with an exclusive cost of 26.33\% of total running time.
+Performance fixes were divised.
 
-Two performance fixes were divised.
+\begin{verbatim}
+ void Body::exert_force_unidirectionally(const Body& there) {
+     const double m1 = m;
+     const double m2 = there.m;
 
-% \begin{lstlisting}
-% \end{lstlisting}
+     const double delta_x = there.x - x;
+     const double delta_y = there.y - y;
 
-Here, we are recalculating
+-    const double r = distance(x, y, there.x, there.y);
++    const double r = hypot(delta_x, delta_y);
+-    const double r2 = r * r;
++    const double r2 = pow(r, 2);
 
-gcc -g -lstdc++ -Wall -pedantic -Wextra -std=c++11 -lm -O3 -march=native Body.o
-QuadTree.o main.o utils.o -o nbody
-Barnes-Hut enabled: false
-Leapfrog enabled: true
-Total CPU time was 13.938825
-12000001 simulation steps computed
+     const double F = (G * m1 * m2) / r2;
+
+     // turn the displacement vector between our two points into a force vector
+     // of the desired magnitude
+     const double scale_factor = F / r;
+
+     Fx += delta_x * scale_factor;
+     Fy += dumpsdelta_y * scale_factor;
+ }
+\end{verbatim}
+
+Instead of recalculating $\Delta x$ and $\Delta y$ twice (the second time in \texttt{distance}), we calculate them only once and make a direct call to \texttt{hypot}, rather than making a call to \texttt{distance}.
+We also used the specialized \texttt{pow} provided by \texttt{cmath}.
+This yielded a recognizable 15\%ish performance improvement.
+
+\begin{verbatim}
+diff --git a/Assignment_1/src/Body.cpp b/Assignment_1/src/Body.cpp
+
++void Body::exert_force_bidirectionally(Body& there) {
++    const double m1 = m;
++    const double m2 = there.m;
++
++    const double delta_x = there.x - x;
++    const double delta_y = there.y - y;
++
++    const double r = hypot(delta_x, delta_y);
++    const double r2 = pow(r, 2);
++
++    const double F = (G * m1 * m2) / r2;
++
++    // turn the displacement vector between our two points into a force vector
++    // of the desired magnitude
++    const double scale_factor = F / r;
++
++    Fx += delta_x * scale_factor;
++    Fy += delta_y * scale_factor;
++
++    there.Fx -= delta_x * scale_factor;
++    there.Fy -= delta_y * scale_factor;
++}
+
+diff --git a/Assignment_1/src/main.cpp b/Assignment_1/src/main.cpp
+@@ -258,8 +259,7 @@ int main(int argc, char **argv) {
+                 for (size_t j = i + 1; j < bodies.size(); j++) {
+                     auto& y = bodies[j];
+-                    x.exert_force_unidirectionally(y);
+-                    y.exert_force_unidirectionally(x);
++                    x.exert_force_bidirectionally(y);
+                 }
+             }
+         }
+\end{verbatim}
+
+This fix halved execution time, for obvious reasons.\footnote{I had to throw out all my old profile data}.
+
+\begin{verbatim}
+diff --git a/Assignment_1/src/Body.cpp b/Assignment_1/src/Body.cpp
+@@ -83,11 +82,11 @@ double Body::kinetic_energy() const {
+ double Body::gravitational_potential_energy(const Body& there) const {
+     const double R = distance(x, y, there.x, there.y); // final distance, aka, to edge
+
+-    return (-G * m * there.m) / R;
++    return (-Gm * there.m) / R;
+ }
+
+@@ -101,7 +100,7 @@ void Body::exert_force_unidirectionally(const Body& there) {
+
+-    const double F = (G * m1 * m2) / r2;
++    const double F = (Gm * m2) / r2;
+
+@@ -113,7 +112,6 @@ void Body::exert_force_unidirectionally(const Body& there) {
+
+-    const double F = (G * m1 * m2) / r2;
++    const double F = (Gm * m2) / r2;
+
+@@ -189,6 +189,7 @@ std::vector<Body> parse_input_file(std::ifstream& input_fh) {
+
+     for (size_t i = 0; i < bodies.size(); i++) {
+         bodies[i].m = masses[i];
++        bodies[i].Gm = G * masses[i];
+     }
+
+\end{verbatim}
+
+Precomputing $Gm$ yielded a 3\%ish performance improvement.
 

### 2018-08-16 Assignment_1/run.sh
Include <memory>

@@ -2,5 +2,5 @@ hash=`git rev-parse --short HEAD`
 name="$1-$hash"
 
 make && ./nbody 6000000 0.016666666666666666 10 "inputs/$1" > "outputs/$name.data" 
-python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
+# python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-16 Assignment_1/src/QuadTree.hpp

@@ -1,6 +1,7 @@
 #ifndef _QuadTree_h
 #define _QuadTree_h
 #include "Body.hpp"
+#include <memory>
 
 class QuadTree {
     public:

### 2018-08-16 Assignment_1/run.sh
Add export notifier

@@ -2,5 +2,6 @@ hash=`git rev-parse --short HEAD`
 name="$1-$hash"
 
 make && ./nbody 6000000 0.016666666666666666 10 "inputs/$1" > "outputs/$name.data" 
+echo "Exported data to ./outputs/$name.data"
 python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-16 Assignment_1/run_example.sh
Add run_example

@@ -1,0 +1,1 @@
+./run.sh in_8_body_counterclockwise

### 2018-08-17 Assignment_1/run.sh
Overhaul commandline parsing

@@ -1,7 +1,14 @@
 hash=`git rev-parse --short HEAD`
 name="$1-$hash"
 
-make && ./nbody 6000000 0.016666666666666666 10 "inputs/$1" > "outputs/$name.data" 
-echo "Exported data to ./outputs/$name.data"
+# numTimeSteps outputInterval deltaT inputFile
+time=10
+fps=60
+
+numTimeSteps=1000000
+outputInterval=`python3 -c  "print(($numTimeSteps / $time) / $fps)"`
+deltaT=`python3 -c  "print($time / $numTimeSteps)"`
+
+make && ./nbody $numTimeSteps $outputInterval $deltaT  "inputs/$1" > "outputs/$name.data" && echo "Exported data to ./outputs/$name.data"
 # python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-17 Assignment_1/src/main.cpp

@@ -105,13 +105,13 @@ void dump_timestep(double timestamp, const std::vector<Body>& bodies) {
 
 void dump_meta_info(
     unsigned int num_time_steps,
-    double output_interval,
+    unsigned int output_interval,
     double delta_t,
     const std::vector<Body>& bodies
 ) {
     const unsigned int bodies_n = bodies.size();
 
-    fprintf(stdout, "%d %d %f %f\n", bodies_n, num_time_steps, output_interval, delta_t); 
+    fprintf(stdout, "%d %d %d %f\n", bodies_n, num_time_steps, output_interval, delta_t); 
 }
 
 /*
@@ -203,14 +203,13 @@ int main(int argc, char **argv) {
     }
 
     const unsigned int num_time_steps = std::stoi(argv[1]);
-    const double output_interval = std::stod(argv[2]);
-    const double delta_t = std::stod(argv[3]);
+    const unsigned int desired_simulation_steps = num_time_steps * (ENABLE_LEAPFROG ? 2 : 1);
 
-    const double timestep = delta_t / num_time_steps;
+    const unsigned int output_interval = std::stoi(argv[2]);
+    const unsigned int output_simulation_step_interval = output_interval * (ENABLE_LEAPFROG ? 2 : 1);
+
+    const double timestep = std::stod(argv[3]);
     const double halfstep = timestep / 2;
-
-    const unsigned int output_step_interval = 
-        output_interval * (ENABLE_LEAPFROG ? 2 : 1) / timestep;
 
     const std::string &input_filename = argv[4];
 
@@ -223,7 +222,7 @@ int main(int argc, char **argv) {
     // TODO: can we have a function that inlines and updates both of these?
     unsigned int step = 0;
 
-    dump_meta_info(num_time_steps, output_interval, delta_t, bodies);
+    dump_meta_info(num_time_steps, output_interval, timestep, bodies);
     dump_masses(bodies);
     dump_timestep(t, bodies);
 
@@ -232,7 +231,7 @@ int main(int argc, char **argv) {
 
     double start = cpu_time();
 
-    while (t < delta_t) {
+    while (step < desired_simulation_steps) {
         QuadTree root;
 
         if (ENABLE_BARNES_HUT && step % 2 == FROG) {
@@ -291,7 +290,7 @@ int main(int argc, char **argv) {
             step++;
         }
 
-        if (step % output_step_interval == 0) {
+        if (step % output_simulation_step_interval == 0) {
             dump_timestep(t, bodies);
         }
     }

### 2018-08-17 Assignment_1/viz.py

@@ -48,8 +48,10 @@ def main():
         contents = f.read()
 
         bodies_n, timestep_n, interval, delta_t = contents.split('\n')[0].split()
-        bodies_n, timestep_n, interval, delta_t = int(bodies_n), int(timestep_n), float(interval), float(delta_t)
+        bodies_n, timestep_n, interval, delta_t = int(bodies_n), int(timestep_n), int(interval), float(delta_t)
         print({"input": in_filename, "numBodies": bodies_n, "numTimeSteps": timestep_n, "outputInterval": interval, "deltaT": delta_t})
+
+        time_per_frame = interval * delta_t
 
         masses = [ float(i) for i in contents.split('\n')[1:bodies_n + 1] ]
         assert(len(masses) == bodies_n)
@@ -88,7 +90,7 @@ def main():
         )
 
         cursor = ax2.axvline(x = 0, animated=True)
-        fps = round(1 / interval)
+        fps = round(1 / time_per_frame)
         frames = np.arange(0, len(timesteps))
         time = np.true_divide(frames, fps)
 
@@ -111,8 +113,9 @@ def main():
 
             return particles, cursor
 
+        print("Constructing an animation with", time_per_frame, "seconds dedicated to each frame")
         ani = animation.FuncAnimation(fig, animate, init_func=init, frames=frames,
-                                    interval=interval * 1000, blit=True)
+                                    interval=time_per_frame * 1000, blit=True)
 
         if len(sys.argv) == 1 + 2:
 
@@ -129,6 +132,7 @@ def main():
             )
             # ani.save('out.gif', writer='imagemagick')
         else:
+            print("Animating at", fps, "FPS")
             plt.show()
 
 if __name__ == "__main__":

### 2018-08-17 Assignment_1/analysis.py
Upload analysis

@@ -1,0 +1,48 @@
+#!/usr/local/bin/python3
+__author__ = "Maxwell Bo (43926871)"
+
+import sys
+import json
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+
+def main():
+    if len(sys.argv) not in [1 + 1, 1 + 2]:
+        print("input [output]")
+        exit(1)
+
+    in_filename = sys.argv[1]
+
+    with open(in_filename, "r") as f:
+        ###########
+        # PARSING #
+        ###########
+        malformed_json = f.read()
+
+        wellformed_json = "[" + malformed_json.replace('\'', '\"') + "]"
+
+        data = json.loads(wellformed_json)
+
+        print(data)
+
+        ##########
+        # LAYOUT #
+        ##########
+        fig = plt.figure(figsize=(6, 6))
+        plt.boxplot([(entry['numBodies'], entry['time']) for entry in data])
+
+        ############
+        # PLOTTING #
+        ############
+
+        if len(sys.argv) == 1 + 2:
+
+            out_filename = sys.argv[2]
+            # ani.save('out.gif', writer='imagemagick')
+        else:
+            plt.show()
+
+if __name__ == "__main__":
+    main()

### 2018-08-17 Assignment_1/report.tex
Update the everything

@@
 
-I personally believe that the simulation is \textit{mostly} right.
-By visualising the results with \texttt{matplotlib}, we see something that resembles an $n$-body simulator\footnote{and that's good enough for me}.
-Distressingly, the total energy of the system is not constant throughout the system.
-When bodies are in close proxmity, anomalous energies that are inconsistent with the energy curve are observed.
-Furthermore, simulations with higher numbers of bodies produce random low energy outliers, but have a smoother total energy curve.
-There seems to be no signfiicant difference between Euler method and Leapfrog, with respect to energy anomalies.
+I personally believe that the simulation is relatively accurate.
+By visualising the results with \texttt{matplotlib}, we see something that resembles an $n$-body simulator.
+The total energy is flat, observing coefficients of variation as low as 0.001\%.
+Euler's method consistently demonstrated coefficients of variation three times higher than that of Leapfrog.
+Due to recommendations by literature, and observed data, the use of Euler's method was gradually phased out during my testing to speed up the process.
+
+Barnes-Hut caused a significant increase in observed coefficient of variation, averaging 0.08\% across multiple runs.
+Furthermore, total energy was observed to step up and down at varying intervals\ref{fig:anomaly}.
+I suspect that this was because certain force calculations were causing groups of bodies to be approximated as a single pseudobody, after other had strayed too far from the pseudobody's quadtree's node's centre of mass.
+
+When bodies are in close proxmity, anomalous total energies are observed\ref{fig:anomaly}.
+Furthermore, simulations with higher numbers of bodies produce more random low energy outliers (likely due to a greater number of close encounters).
+There seems to be no signficant difference between Euler method and Leapfrog, with respect to observation of anomalies.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{energy_anomaly}
+\label{fig:anomaly}
+\includegraphics[width=\textwidth]{energy_anomaly}
 \end{figure}
@@
 \begin{figure}[b]
-\caption{Observed energy anomaly while bodies in close proxmity - Euler method}
+\caption{Barnes-Hut energy variation - Leapfrog}
 \centering
-\includegraphics[width=0.5\textwidth]{euler_energy_anomaly}
-\end{figure}
-
-\begin{figure}[b]
-\caption{Observed energy anomaly while bodies in close proxmity - Euler method}
-\centering
-\includegraphics[width=0.5\textwidth]{euler_relaxed_energy_anomaly}
-\end{figure}
-
-\begin{figure}[b]
-\caption{Observed energy anomaly - random outlier - Leapfrog}
-\centering
-\includegraphics[width=0.5\textwidth]{relaxed_energy_anomaly}
+\label{fig:barnes_hut}
+\includegraphics[width=\textwidth]{barnes_hut}
 \end{figure}

### 2018-08-17 Assignment_1/analysis.py

@@ -25,22 +25,36 @@ def main():
 
         data = json.loads(wellformed_json)
 
-        print(data)
-
         ##########
         # LAYOUT #
         ##########
-        fig = plt.figure(figsize=(6, 6))
-        plt.boxplot([(entry['numBodies'], entry['time']) for entry in data])
+        fig = plt.figure(figsize=(6, 8))
+        gs = gridspec.GridSpec(2, 1) 
+
+        ax1 = plt.subplot(gs[0])
+        ax1.set_title("60000 timestep running time")
+        ax1.set(xlabel="x (m)", ylabel="y (m)")
+
+        ax2 = plt.subplot(gs[1])
+        ax2.set_title("60000 timestep running time with Barnes-Hut enabled")
+        ax2.set(xlabel="t (s)", ylabel="E (J)")
+
+        plt.tight_layout()
+
+        xs = [(entry['numBodies']) for entry in data if not entry['barnesHut']]
+        ys = [(entry['time']) for entry in data if not entry['barnesHut']]
+        ax1.plot(xs, ys, "x")
+
+        xs = [(entry['numBodies']) for entry in data if entry['barnesHut']]
+        ys = [(entry['time']) for entry in data if entry['barnesHut']]
+        ax2.plot(xs, ys, "x")
 
         ############
         # PLOTTING #
         ############
 
         if len(sys.argv) == 1 + 2:
-
             out_filename = sys.argv[2]
-            # ani.save('out.gif', writer='imagemagick')
         else:
             plt.show()
 

### 2018-08-17 Assignment_1/run.sh

@@ -10,5 +10,5 @@ outputInterval=`python3 -c  "print(($numTimeSteps / $time) / $fps)"`
 deltaT=`python3 -c  "print($time / $numTimeSteps)"`
 
 make && ./nbody $numTimeSteps $outputInterval $deltaT  "inputs/$1" > "outputs/$name.data" && echo "Exported data to ./outputs/$name.data"
-# python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
+python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-08-17 Assignment_1/src/main.cpp

@@ -62,7 +62,6 @@ double calculate_total_energy(const std::vector<Body>& bodies) {
             auto& y = bodies[j];
 
             acc += x.gravitational_potential_energy(y);
-            acc += y.gravitational_potential_energy(x);
         }
     }
     
@@ -234,20 +233,24 @@ int main(int argc, char **argv) {
     while (step < desired_simulation_steps) {
         QuadTree root;
 
-        if (ENABLE_BARNES_HUT && step % 2 == FROG) {
-            const double root_x = 0;
-            const double root_y = 0;
-            const double radius = maximum_deviation_from_root(bodies) + 1;
-            // the quad-tree uses half the width as an implementation detail
-            // called "radius". We're trying to make a QuadTree that encapsulates
-            // the most distant body
+        // only the Frog step (not the Leap step) needs updated forces
+        // hence all the (step % 2 == FROG) tests
+        const bool need_force_calc = !ENABLE_LEAPFROG || step % 2 == FROG;
 
-            root = QuadTree(root_x, root_y, radius);
+        if (need_force_calc) {
+            if (ENABLE_BARNES_HUT) {
+                const double root_x = 0;
+                const double root_y = 0;
+                const double radius = maximum_deviation_from_root(bodies) + 1;
+                // the quad-tree uses half the width as an implementation detail
+                // called "radius". We're trying to make a QuadTree that encapsulates
+                // the most distant body
 
-            assert(root.insert_all(bodies));
-        }
+                root = QuadTree(root_x, root_y, radius);
 
-        if (!ENABLE_LEAPFROG || step % 2 == FROG) {
+                assert(root.insert_all(bodies));
+            }
+
             for (auto& body: bodies) {
                 body.reset_force();
 
@@ -255,15 +258,15 @@ int main(int argc, char **argv) {
                     root.calculate_force(body);
                 }
             }
-        }
 
-        if (!ENABLE_BARNES_HUT) {
-            for (size_t i = 0; i < bodies.size() - 1; i++) {
-                auto& x = bodies[i];
+            if (!ENABLE_BARNES_HUT) {
+                for (size_t i = 0; i < bodies.size() - 1; i++) {
+                    auto& x = bodies[i];
 
-                for (size_t j = i + 1; j < bodies.size(); j++) {
-                    auto& y = bodies[j];
-                    x.exert_force_bidirectionally(y);
+                    for (size_t j = i + 1; j < bodies.size(); j++) {
+                        auto& y = bodies[j];
[…]

### 2018-08-17 Assignment_1/viz.py

@@ -3,6 +3,7 @@ __author__ = "Maxwell Bo (43926871)"
 
 import sys
 import math
+import statistics
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.animation as animation
@@ -94,7 +95,10 @@ def main():
         frames = np.arange(0, len(timesteps))
         time = np.true_divide(frames, fps)
 
-        ax2.plot(time, [timestep.total_energy for timestep in timesteps], "bo", markersize=2)
+        energies = [timestep.total_energy for timestep in timesteps]
+        print("Coefficient of variation of energy is", statistics.stdev(energies) / statistics.mean(energies) * 100)
+
+        ax2.plot(time, energies, "bo", markersize=2)
 
         def init():
             return particles, cursor

### 2018-08-17 Assignment_1/viz.py
Update viz and log

@@ -125,7 +125,7 @@ def main():
 
             out_filename = sys.argv[2]
 
-            print("Exporting at", fps, "FPS")
+            print("Exporting to", out_filename, "at", fps, "FPS")
             ani.save(out_filename, writer=animation.FFMpegWriter(
                 fps=fps, 
                 metadata={

### 2018-08-17 Assignment_1/report.tex

@@
 Dehen and Read note that the Euler method `performs very poorly in practice', further noting that `errors are proportional to $\Delta t^2$'.
-They contrast it with the second-order leapfrog symplectic integrator, which is `heavily used in collisionless N-body applications'.\cite{dehen11a}
+They contrast it with the second-order \textit{Leapfrog} symplectic integrator, which is `heavily used in collisionless N-body applications'.\cite{dehen11a}
 
-Leapfrog can be expressed many in forms\cite{Skeel1993} including a synchronised form:
+\textit{Leapfrog} can be expressed many in forms\cite{Skeel1993} including a synchronised form:
 
@@
     \item Inserted all \texttt{Body}s into a fresh \texttt{QuadTree} on full timesteps, traversing the \texttt{QuadTree} with every \texttt{Body} to calculate forces ($O(n \log n)$)
-    \item Performed the appropriate Leapfrog step to derive new velocities \textit{or} positions
+    \item Performed the appropriate \textit{Leapfrog} step to derive new velocities \textit{or} positions
     \item Output all necessary data
@@
 The total energy is flat, observing coefficients of variation as low as 0.001\%.
-Euler's method consistently demonstrated coefficients of variation three times higher than that of Leapfrog.
+Euler's method consistently demonstrated coefficients of variation three times higher than that of \textit{Leapfrog}.
 Due to recommendations by literature, and observed data, the use of Euler's method was gradually phased out during my testing to speed up the process.
 
-Barnes-Hut caused a significant increase in observed coefficient of variation, averaging 0.08\% across multiple runs.
+\textit{Barnes-Hut} caused a significant increase in observed coefficient of variation, averaging 0.08\% across multiple runs.
 Furthermore, total energy was observed to step up and down at varying intervals\ref{fig:anomaly}.
@@
 Furthermore, simulations with higher numbers of bodies produce more random low energy outliers (likely due to a greater number of close encounters).
-There seems to be no signficant difference between Euler method and Leapfrog, with respect to observation of anomalies.
+There seems to be no signficant difference between Euler method and \textit{Leapfrog}, with respect to observation of anomalies.
 
@@
 Henceforth, the use of the `recognizable' refers to eyeballing the output data, and making no significant effort to investigate the data more rigorously.
+
+The \textit{Barnes-Hut} algorithm appeared to be dominated by its constant factor.\ref{fig:performance} Generating input files with a high number of bodies that did not cause 'spontaneous combustion' (where bodies would fly away from each other in every direction), and unconstrained Quadtree growth (where bodies achieve escape velocity, rapidly expanding the size of the tree and causing a collapse in simulation accuracy) proved difficult.
+Further testing on \texttt{goliath} is needed to discover the $n$ that causes \textit{Barnes-Hut} to be more performant than the brute-force approach.
+
+\begin{figure}[b]
+\caption{Scaling characteristics}
+\centering
+\label{fig:performance}
+\includegraphics[width=\textwidth]{performance}
+\end{figure}
 

### 2018-08-17 Assignment_1/analysis.py
Update analysis

@@ -33,11 +33,11 @@ def main():
 
         ax1 = plt.subplot(gs[0])
         ax1.set_title("60000 timestep running time")
-        ax1.set(xlabel="x (m)", ylabel="y (m)")
+        ax1.set(ylabel="time elapsed (s)", xlabel="number of bodies (n)")
 
         ax2 = plt.subplot(gs[1])
         ax2.set_title("60000 timestep running time with Barnes-Hut enabled")
-        ax2.set(xlabel="t (s)", ylabel="E (J)")
+        ax2.set(ylabel="time elapsed (s)", xlabel="number of bodies (n)")
 
         plt.tight_layout()
 

### 2018-08-17 Assignment_1/analysis.py
Update everything

@@ -33,11 +33,11 @@ def main():
 
         ax1 = plt.subplot(gs[0])
         ax1.set_title("60000 timestep running time")
-        ax1.set(ylabel="time elapsed (s)", xlabel="number of bodies (n)")
+        ax1.set(ylabel="time elapsed (s)", xlabel="bodies (n)")
 
         ax2 = plt.subplot(gs[1])
         ax2.set_title("60000 timestep running time with Barnes-Hut enabled")
-        ax2.set(ylabel="time elapsed (s)", xlabel="number of bodies (n)")
+        ax2.set(ylabel="time elapsed (s)", xlabel="bodies (n)")
 
         plt.tight_layout()
 

### 2018-09-19 run.sh
Init

@@ -10,5 +10,5 @@ outputInterval=`python3 -c  "print(($numTimeSteps / $time) / $fps)"`
 deltaT=`python3 -c  "print($time / $numTimeSteps)"`
 
 make && ./nbody $numTimeSteps $outputInterval $deltaT  "inputs/$1" > "outputs/$name.data" && echo "Exported data to ./outputs/$name.data"
-python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
+# python3 viz.py "outputs/$name.data" "outputs/$name.mp4" && open "outputs/$name.mp4"
 

### 2018-09-19 src/main.cpp

@@ -6,6 +6,7 @@
 #include <time.h>
 #include <algorithm>
 #include <limits>
+#include <omp.h>
 
 // #include "CImg.h"
 #include "Body.hpp"
@@ -29,6 +30,8 @@ double maximum_deviation_from_root(const std::vector<Body>& bodies) {
     double lowest_y = std::numeric_limits<double>::max();
     double highest_y = std::numeric_limits<double>::min();
 
+    // CANDIDATE
+    // #pragma omp parallel default(shared)
     for (const auto& body: bodies) {
         if (body.x < lowest_x) {
             lowest_x = body.x;
@@ -51,6 +54,7 @@ double maximum_deviation_from_root(const std::vector<Body>& bodies) {
 double calculate_total_energy(const std::vector<Body>& bodies) {
     double acc = 0;
 
+    // CANDIDATE
     for (auto& body : bodies) {
         acc += body.kinetic_energy();
     }
@@ -58,6 +62,8 @@ double calculate_total_energy(const std::vector<Body>& bodies) {
     for (size_t i = 0; i < bodies.size() - 1; i++) {
         auto& x = bodies[i];
 
+        // CANDIDATE
+        // #pragma omp for reduction(+:acc)
         for (size_t j = i + 1; j < bodies.size(); j++) {
             auto& y = bodies[j];
 
@@ -245,38 +251,36 @@ int main(int argc, char **argv) {
             }
 
             if (!ENABLE_BARNES_HUT) {
-                for (size_t i = 0; i < bodies.size() - 1; i++) {
-                    auto& x = bodies[i];
+                #pragma omp parallel
+                {
+                    #pragma omp for
+                    for (size_t i = 0; i < bodies.size() - 1; i++) {
+                        auto& x = bodies[i];
 
-                    for (size_t j = i + 1; j < bodies.size(); j++) {
-                        auto& y = bodies[j];
-                        x.exert_force_bidirectionally(y);
+                        for (size_t j = i + 1; j < bodies.size(); j++) {
+                            auto& y = bodies[j];
+
+                            #pragma omp critical
+                            {
+                                x.exert_force_bidirectionally(y);
+                            }
+                        }
                     }
                 }
             }
         }
 
-        if (ENABLE_LEAPFROG) {
-            for (auto& body: bodies) {
-                if (step % 2 == LEAP) {
-                    body.leap(timestep);
-                }
-                else {
-                    body.frog(timestep);
-                }
+        for (auto& body: bodies) {
+            if (step % 2 == LEAP) {
[…]

### 2018-09-21 src/Body.cpp
Body

@@ -36,11 +36,13 @@ void Body::euler_integrate(double dt) {
     y += dy;
 }
 
+// pos and velocity
 void Body::leap(double dt) {
     x = x + (vx * dt);
     y = y + (vy * dt);
 }
 
+// velocity, mass, and force
 void Body::frog(double dt) {
     const double ax = Fx / m;
     const double ay = Fy / m;

### 2018-09-21 src/main.cpp

@@ -20,6 +20,8 @@ const bool ENABLE_BARNES_HUT = false;
 const bool ENABLE_LEAPFROG = true;
 const bool ENABLE_LOGGING = true;
 
+const int root = 0;
+
 double cpu_time(void) {
     return (double)clock() / (double)CLOCKS_PER_SEC;
 }
@@ -132,9 +134,8 @@ void dump_meta_info(
 }
 
 
-void broadcast_bodies(std::vector<Body> bodies, int rank) {
+void broadcast_bodies(std::vector<Body> bodies, int size, int rank) {
     size_t bodies_n = bodies.size();
-    int broadcast_root = 0;
     
     double x[bodies_n];
     double y[bodies_n];
@@ -143,7 +144,7 @@ void broadcast_bodies(std::vector<Body> bodies, int rank) {
 
     // This essentially round-trips the data for rank == 0 (which is good)
     // meaning it gets copied to all other nodes
-    if (rank == broadcast_root) {
+    if (rank == root) {
         for (size_t i = 0; i < bodies.size(); i++) {
             auto& body = bodies[i];
             x[i] = body.x;
@@ -153,12 +154,12 @@ void broadcast_bodies(std::vector<Body> bodies, int rank) {
         }
     }
 
-    MPI_Bcast(x, bodies_n, MPI_DOUBLE, broadcast_root, MPI_COMM_WORLD);
-    MPI_Bcast(y, bodies_n, MPI_DOUBLE, broadcast_root, MPI_COMM_WORLD);
-    MPI_Bcast(vx, bodies_n, MPI_DOUBLE, broadcast_root, MPI_COMM_WORLD);
-    MPI_Bcast(vy, bodies_n, MPI_DOUBLE, broadcast_root, MPI_COMM_WORLD);
+    MPI_Bcast(x, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
+    MPI_Bcast(y, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
+    MPI_Bcast(vx, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
+    MPI_Bcast(vy, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
 
-    if (rank != broadcast_root) {
+    if (rank != root) {
         for (size_t i = 0; i < bodies.size(); i++) {
             auto& body = bodies[i];
             body.x = x[i];
[…]

### 2018-09-21 src/main.cpp
Remove broadcast

@@ -133,43 +133,6 @@ void dump_meta_info(
     fprintf(stdout, "%d %d %d %f\n", bodies_n, num_time_steps, output_interval, delta_t); 
 }
 
-
-void broadcast_bodies(std::vector<Body> bodies, int size, int rank) {
-    size_t bodies_n = bodies.size();
-    
-    double x[bodies_n];
-    double y[bodies_n];
-    double vx[bodies_n];
-    double vy[bodies_n];
-
-    // This essentially round-trips the data for rank == 0 (which is good)
-    // meaning it gets copied to all other nodes
-    if (rank == root) {
-        for (size_t i = 0; i < bodies.size(); i++) {
-            auto& body = bodies[i];
-            x[i] = body.x;
-            y[i] = body.y;
-            vx[i] = body.vx;
-            vy[i] = body.vy;
-        }
-    }
-
-    MPI_Bcast(x, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
-    MPI_Bcast(y, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
-    MPI_Bcast(vx, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
-    MPI_Bcast(vy, bodies_n, MPI_DOUBLE, root, MPI_COMM_WORLD);
-
-    if (rank != root) {
-        for (size_t i = 0; i < bodies.size(); i++) {
-            auto& body = bodies[i];
-            body.x = x[i];
-            body.y = y[i];
-            body.vx = vx[i];
-            body.vy = vy[i];
-        }
-    }
-}
-
 std::vector<Body> scatter_bodies(std::vector<Body> bodies, int size, int rank) {
     unsigned int bodies_n = bodies.size();
 
@@ -512,11 +475,12 @@ int main(int argc, char **argv) {
             }
         }
 
-        std::vector<Body> sbodies = scatter_bodies(bodies, size, rank);
+        // std::vector<Body> sbodies = scatter_bodies(bodies, size, rank);
+        // std::vector<Body bodies = sbodies;
 
         #pragma omp parallel for shared(bodies)
-        for (size_t i = 0; i < sbodies.size(); i++) {
-            auto& body = sbodies[i];
+        for (size_t i = 0; i < bodies.size(); i++) {
+            auto& body = bodies[i];
 
             if (step % 2 == LEAP) {
                 body.leap(timestep);
[…]

### 2018-09-21 go.sh
Add go.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#SBATCH --partition=coursework
+#SBATCH --job-name=mbody
+#SBATCH --nodes=4
+#SBATCH --ntasks=4
+#SBATCH --ntasks-per-node=1
+#SBATCH --cpus-per-task=4
+#SBATCH --error=stderr.log
+DATE=$(date +"%Y%m%d%H%M")
+echo "time started "$DATE
+export OMP_NUM_THREADS=${SLURM_CPUS_PER_TASK}
+echo "This is job ’$SLURM_JOB_NAME’ (id: $SLURM_JOB_ID) running on the following nodes:"
+echo $SLURM_NODELIST
+echo "running with OMP_NUM_THREADS=$OMP_NUM_THREADS "
+echo "running with SLURM_TASKS_PER_NODE=$SLURM_TASKS_PER_NODE "
+echo "--------------------:"
+module load mpi
+time mpirun -n 4 ./nbody 5 1 1 inputs/in_10000

### 2018-09-21 reports/ms1_report.tex
Init
Big refactor

--- 2018-08-17 Assignment_1/report.tex
@@
 
+Strangely, $n = 16$ with \textit{Barnes-Hut} enabled took longer than expected, when compared to $n \in \{ 4, 8, 12 \}$. Further investigation is required.
 \begin{figure}[b]
@@
 
+\section*{Parallelism}
+A description of your parallel implementation.
+How do you make use of OpenMP and MPI?
+How is information and work shared between threads and processes.
+Any optimisations applied and/or challenges encountered will go here.
+\subsection*{Implementation}
+\subsubsection*{OpenMP}
+
+I slapped as many \texttt{\#pragma omp parallel for shared(bodies)} on as many loops as I could find, and quickly discovered massive lock contention issues associated with \texttt{\#pragma omp critical}.
+Pairwise force updates between bodies:
+
+\begin{verbatim}
+    for (size_t i = 0; i < bodies.size() - 1; i++) {
+        auto& x = bodies[i];
+
+        for (size_t j = i + 1; j < bodies.size(); j++) {
+            auto& y = bodies[j];
+                x.exert_force_bidirectionally(y);
+        }
+    }
+\end{verbatim}
+
+where
+
+\begin{verbatim}
+void Body::exert_force_bidirectionally(Body& there) {
+    ...
+    #pragma omp critical
+    {
+    Fx += delta_x * scale_factor;
+    Fy += delta_y * scale_factor;
+
+    there.Fx -= delta_x * scale_factor;
+    there.Fy -= delta_y * scale_factor;
+    }
+}
+\end{verbatim}
+
+which had provided massive performance gains in the serial implementation over unidirectional updates was now causing massive parallel slowdowns, where increasing the \texttt{OMP\_NUM\_THREADS} caused a proportional increase in walltime.
+Instead, a \texttt{critical}-less comparison schema was devised,
+
+\begin{verbatim}
+#pragma omp parallel for shared(bodies)
+for (size_t i = 0; i < bodies.size(); i++) {
+    auto& x = bodies[i];
+
+    for (size_t j = 0; j < bodies.size(); j++) {
+        auto& y = bodies[j];
+
+        if (&bodies[i] != &bodies[j]) {
+            // XXX: Do NOT swap these around. You will cause
+            // race conditions
+            x.exert_force_unidirectionally(y);
+        }
+    }
+}
+\end{verbatim}
+
+that provided the sought after parallel speedup.
+This parallel speedup was found to be present at where $n$-bodies was high.
+Testing focussed primarily on $n > 5000$,\footnote{In MS1, I failed to test simulations with a large number of $n$, as I was focussed on testing the accuracy of the implementation, and generated input files would cause high numbers of collisions that made accuracy assessment difficult.
+Testing $n=10000$, we find a 10x speedup with by using a serial Barnes-Hut implementation (Avg.
+$1.12$s, as opposed to $9.87$s without Barnes-Hut).
+Barnes-Hut afforded a much simpler parallelization scheme, as each body could calculate its force against the quadtree concurrently, without locks} where force and integration calculation overhead dominated thread creation overhead.
+Parallel speedup vanished around the $n \approx 500$ mark.
+
+% Increasing \texttt{OMP_NUM_THREADS} beyond \texttt{4} failed to provide a speedup at $n = 10000$.
+
+\subsubsection*{MPI}
+
+Looking at our high-level implementation, the following two steps consume the vast majority of our computing power:
+
+\begin{enumerate}
+    \item Inserted all \texttt{Body}s into a fresh \texttt{QuadTree} on full timesteps, traversing the \texttt{QuadTree} with every \texttt{Body} to calculate forces ($O(n \log n)$)
+    \item Performed the appropriate \textit{Leapfrog} step to derive new velocities \textit{or} positions
+\end{enumerate}
+
+The \textit{Leapfrog} step is the easiest step to distribute amongst many nodes, as it does not require the comparison of a body to all other bodies - each node can compute the step on a subset of all bodies, and later combine each subset with relatively simple MPI code.
+
+This can be achieved with \texttt{MPI\_Scatterv} and \texttt{MPI\_Gatherv}
+\footnote{These \texttt{v} variants support custom chunking sizes e.g.
+if we have 15 bodies we can send $4, 4, 4, 3$ bodies to each node respectively, which \texttt{MPI\_Scatter} and \texttt{MPI\_Gather} do not support dynamic chunking}.
+We decompose each dimension of our \texttt{Body}s into arrays, (e.g.
+the $x$ position, the $x$ veolocity, etc.), \texttt{MPI\_Scatterv} these arrays, recompose them into \texttt{Body}s, run \texttt{Leapfrog} on this subset on each node, decompose the subset into arrays, \texttt{MPI\_Gatherv} these arrays back to the root node, and recompose them into $\texttt{Bodys}$.
+
+\begin{verbatim}
+std::vector<Body> sbodies = scatter_bodies(bodies, size, rank);
+
+#pragma omp parallel for shared(bodies)
+for (size_t i = 0; i < sbodies.size(); i++) {
+    auto& body = sbodies[i];
+
+    if (step % 2 == LEAP) {
+        body.leap(timestep);
+    }
+    else {
+        body.frog(timestep);
+    }
+}
+
+bodies = gather_bodies(sbodies, size, rank, bodies.size());
+\end{verbatim}
+
+where \texttt{sbodies} is our subset of bodies distributed to each node.
+
+Because force calculation is only performed on the root node, this messaging architecture is \textit{master-slave}, where the root node coordinates gathers and scatters.
+Slave nodes, are, unfortunately, underutilized.
+Further work could be sent distributing force calculation work, doing away with a root node entirely, except for logging.
+
+\includegraphics[width=3cm]{mpi_gather}
+\includegraphics[width=3cm]{mpi_scatter}
+
+% A description of how you have verified the correctness of your parallel implementation. (See comments for MS1)
+\subsection*{Verification \& Correctness}
+
+Because of my confidence in the correctness of the serial implementation, it served as a reference implementation for the verification of the parallel implementation.
+
+Both implementations were run with identical input parameters, and their outputs were \texttt{diff}d.
+When run without MPI enabled, outputs were identical.
+When run with MPI, floating point abnormalities (initially affecting the least significant digits of each floating point value) gradually caused each diff to diverge.
+I hypothesize that the MPI serialization and deserialization procedure was truncating precision and causing this divergence.
+Perhaps some option exists to ensure that MPI guarantees floating point precision.
+
+% A credible plan of jobs to run to measure the scalability of your parallel implementation.
+The idea here is that you could take this plan an start performing timing experiments.
+You should not organise your testing so that if your test plan is curtailed, then you can still draw sensible conclusions from the results you do get.
+\subsection*{Scalability}
+
+We have mulitple dimensions that we can change to measure the scalability of the parallel implementation.
+
+\begin{enumerate}
+    \item The number of bodies ($n$) (\texttt{512, 1024, 2048, 4096, 8192, 16384})
+    \item \texttt{ENABLE\_BARNES\_HUT} (\texttt{true, false})
+    \item \texttt{--nodes} (\texttt{1, 4, 16})
+    \item \texttt{--ntasks} (\texttt{1, 4, 16})
+    \item \texttt{--ntasks-per-node} (\texttt{1, 4})
+    \item \texttt{--cpus-per-task} (\texttt{1, 4, 16})
+\end{enumerate}
+
+Running tests with all valid combinations of these variables, with a \texttt{numTimeSteps} as low as 8, would produce useful data, across multiple dimension, with minimal load on cluster.
+It would be unlikely that this plan would be curtailed.
+
 \medskip

### 2018-09-21 src/main.cpp

@@ -402,7 +402,6 @@ int main(int argc, char **argv) {
 
     std::vector<Body> bodies = parse_input_file(input_fh);
 
-    double start = cpu_time();
     double t = 0; 
     unsigned int step = 0;
 
@@ -416,65 +415,63 @@ int main(int argc, char **argv) {
         fprintf(stderr, "OMP Max threads: %d\n", omp_get_max_threads());
     }
 
+    double start = cpu_time();
     while (step < desired_simulation_steps) {
-        if (rank == root) {
-            QuadTree qroot;
+        QuadTree qroot;
 
-            // only the Frog step (not the Leap step) needs updated forces
-            // hence all the (step % 2 == FROG) tests
-            const bool need_force_calc = step % 2 == FROG;
+        // only the Frog step (not the Leap step) needs updated forces
+        // hence all the (step % 2 == FROG) tests
+        const bool need_force_calc = (rank == root) || step % 2 == FROG;
 
-            if (need_force_calc) {
+        if (need_force_calc) {
+            if (ENABLE_BARNES_HUT) {
+                const double root_x = 0;
+                const double root_y = 0;
+                const double radius = maximum_deviation_from_root(bodies) + 1;
+                // the quad-tree uses half the width as an implementation detail
+                // called "radius". We're trying to make a QuadTree that encapsulates
+                // the most distant body
+
+                qroot = QuadTree(root_x, root_y, radius);
+
+                assert(qroot.insert_all(bodies));
+            }
+
+            #pragma omp parallel for shared(bodies)
+            for (size_t i = 0; i < bodies.size(); i++) {
+                auto& body = bodies[i];
+                body.reset_force();
+
                 if (ENABLE_BARNES_HUT) {
-                    const double root_x = 0;
-                    const double root_y = 0;
-                    const double radius = maximum_deviation_from_root(bodies) + 1;
-                    // the quad-tree uses half the width as an implementation detail
-                    // called "radius". We're trying to make a QuadTree that encapsulates
-                    // the most distant body
-
-                    qroot = QuadTree(root_x, root_y, radius);
-
-                    assert(qroot.insert_all(bodies));
+                    qroot.calculate_force(body);
                 }
+            }
 
+            if (!ENABLE_BARNES_HUT) {
                 #pragma omp parallel for shared(bodies)
                 for (size_t i = 0; i < bodies.size(); i++) {
-                    auto& body = bodies[i];
-                    body.reset_force();
[…]

### 2018-09-21 src/main.cpp
Hmm

@@ -472,12 +472,14 @@ int main(int argc, char **argv) {
                 }
             }
         }
-        // std::vector<Body> sbodies = scatter_bodies(bodies, size, rank);
-        // std::vector<Body bodies = sbodies;
 
-        #pragma omp parallel for shared(bodies)
-        for (size_t i = 0; i < bodies.size(); i++) {
-            auto& body = bodies[i];
+        
+        // std::vector<Body> sbodies = scatter_bodies(bodies, size, rank);
+        std::vector<Body> sbodies = bodies;
+
+        #pragma omp parallel for shared(sbodies)
+        for (size_t i = 0; i < sbodies.size(); i++) {
+            auto& body = sbodies[i];
 
             if (step % 2 == LEAP) {
                 body.leap(timestep);

### 2018-10-16 go.sh
Fix debug instrumentation

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH --partition=smp
 #SBATCH --job-name=mbody
-#SBATCH --nodes=1
-#SBATCH --ntasks=1
-#SBATCH --cpus-per-task=1
+#SBATCH --nodes=4
+#SBATCH --ntasks=4
+#SBATCH --cpus-per-task=4
 #SBATCH --error=stderr.log
 
 DATE=$(date +"%Y%m%d%H%M")

### 2018-10-16 src/main.cpp

@@ -21,7 +21,8 @@ const unsigned int FROG = 1;
 
 const bool ENABLE_BARNES_HUT = false;
 const bool ENABLE_LEAPFROG = true;
-const bool ENABLE_LOGGING = true;
+const bool ENABLE_LOGGING = false;
+const bool ENABLE_DEBUG_LOGGING = true;
 
 const int root = 0;
 
@@ -138,10 +139,17 @@ void dump_meta_info(
 
 void scatter_bodies(const std::vector<Body>& bodies, std::vector<Body>& sbodies, int rank, 
     const std::vector<int>& send_counts, const std::vector<int>& displacements) {
+
     unsigned int receive_count = send_counts[rank];
     unsigned int displacement = displacements[rank];
 
-    fprintf(stderr, "Sent %d from displacement %d from rank%d\n", receive_count, displacement, rank);
+    if (ENABLE_DEBUG_LOGGING) {
+        if (rank == 0) {
+            fprintf(stderr, "[%d] Scattering %d from displacement %d\n", rank, receive_count, displacement);
+        } else {
+            fprintf(stderr, "[%d] Receiving %d from scatter into displacement %d\n", rank, receive_count, displacement);
+        }
+    }
 
     MPI_Scatterv(
         bodies.data(), // sendbuf
@@ -162,7 +170,13 @@ void gather_bodies(std::vector<Body>& bodies, const std::vector<Body>& sbodies, 
     unsigned int send_count = receive_counts[rank];
     unsigned int displacement = displacements[rank];
 
-    fprintf(stderr, "Gathering %d into displacement %d [from rank%d]\n", send_count, displacement, rank);
+    if (ENABLE_DEBUG_LOGGING) {
+        if (rank == 0) {
+            fprintf(stderr, "[%d] Gathering %d into displacement %d\n", rank, send_count, displacement);
+        } else {
+            fprintf(stderr, "[%d] Sending %d for gather from displacement %d\n", rank, send_count, displacement);
+        }
+    }
 
     MPI_Gatherv(
         sbodies.data(), // sendbuf
@@ -247,7 +261,7 @@ std::vector<Body> parse_input_file(std::ifstream& input_fh) {
 
 int main(int argc, char **argv) {
     if (argc != 5) {
-        fprintf(stderr, "numTimeSteps outputInterval deltaT inputFile\n");
+        fprintf(stdout, "numTimeSteps outputInterval deltaT inputFile\n");
[…]

### 2018-10-16 reports/ms1_report.tex
Update report

@@
 \section*{Parallelism}
-A description of your parallel implementation.
+% A description of your parallel implementation.
 How do you make use of OpenMP and MPI?
@@
 
+\includegraphics[width=7cm]{mpi_scatter}
+\includegraphics[width=7cm]{mpi_gather}
+
 \begin{verbatim}
@@
 
-\includegraphics[width=3cm]{mpi_gather}
-\includegraphics[width=3cm]{mpi_scatter}
+\begin{verbatim}
+Total CPU time on rank 0 was 89.820000
+Total CPU time on rank 2 was 36.750000
+Total CPU time on rank 3 was 36.960000
+Total CPU time on rank 1 was 37.040000
+\end{verbatim}
 
@@
     \item \texttt{ENABLE\_BARNES\_HUT} (\texttt{true, false})
-    \item \texttt{--nodes} (\texttt{1, 4, 16})
+    \item \texttt{--nodes} (\texttt{Disabled, 1, 4, 16})
     \item \texttt{--ntasks} (\texttt{1, 4, 16})
@@
     \item \texttt{--cpus-per-task} (\texttt{1, 4, 16})
+    \item \texttt{\#pragma omp parallel schedule(static, dynamic, guided)}
 \end{enumerate}

### 2018-10-16 go.sh
Fix broadcast issues

@@ -1,9 +1,9 @@
 #!/bin/bash
-#SBATCH --partition=smp
+#SBATCH --partition=cosc
 #SBATCH --job-name=mbody
-#SBATCH --nodes=4
-#SBATCH --ntasks=4
-#SBATCH --cpus-per-task=4
+#SBATCH --nodes=1
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=16
 #SBATCH --error=stderr.log
 
 DATE=$(date +"%Y%m%d%H%M")
@@ -14,4 +14,4 @@ echo $SLURM_NODELIST
 echo "running with OMP_NUM_THREADS=$OMP_NUM_THREADS "
 echo "running with SLURM_TASKS_PER_NODE=$SLURM_TASKS_PER_NODE "
 echo "--------------------:"
-time mpirun -n ${SLURM_NPROCS} ./nbody 5 1 1 inputs/in_12_body_counterclockwise
+time mpirun -n ${SLURM_NPROCS} ./nbody 1000000 1666.6666666666667 1e-05 inputs/in_8_body_counterclockwise

### 2018-10-16 src/main.cpp

@@ -141,15 +141,6 @@ void scatter_bodies(const std::vector<Body>& bodies, std::vector<Body>& sbodies,
     const std::vector<int>& send_counts, const std::vector<int>& displacements) {
 
     unsigned int receive_count = send_counts[rank];
-    unsigned int displacement = displacements[rank];
-
-    if (ENABLE_DEBUG_LOGGING) {
-        if (rank == 0) {
-            fprintf(stderr, "[%d] Scattering %d from displacement %d\n", rank, receive_count, displacement);
-        } else {
-            fprintf(stderr, "[%d] Receiving %d from scatter into displacement %d\n", rank, receive_count, displacement);
-        }
-    }
 
     MPI_Scatterv(
         bodies.data(), // sendbuf
@@ -168,15 +159,6 @@ void gather_bodies(std::vector<Body>& bodies, const std::vector<Body>& sbodies, 
     const std::vector<int>& receive_counts, const std::vector<int>& displacements) {
 
     unsigned int send_count = receive_counts[rank];
-    unsigned int displacement = displacements[rank];
-
-    if (ENABLE_DEBUG_LOGGING) {
-        if (rank == 0) {
-            fprintf(stderr, "[%d] Gathering %d into displacement %d\n", rank, send_count, displacement);
-        } else {
-            fprintf(stderr, "[%d] Sending %d for gather from displacement %d\n", rank, send_count, displacement);
-        }
-    }
 
     MPI_Gatherv(
         sbodies.data(), // sendbuf
@@ -273,7 +255,7 @@ int main(int argc, char **argv) {
     MPI_Init(&argc, &argv);
     MPI_Comm_size(comm, &size); // Get the number of processes
     MPI_Comm_rank(comm, &rank); // Get the rank of the process
-    MPI_Type_contiguous(7, MPI_DOUBLE, &MPI_Body); // 7 doubles in the Body struct
+    MPI_Type_contiguous(8, MPI_DOUBLE, &MPI_Body); // 8 doubles in the Body struct
     MPI_Type_commit(&MPI_Body);
 
     // ---------------------------------------------------------------------//
@@ -326,8 +308,7 @@ int main(int argc, char **argv) {
     }
     fprintf(stderr, "\n");
 
-    std::vector<Body> sbodies = {};
-    sbodies.reserve(send_counts[rank]);
+    std::vector<Body> sbodies(send_counts[rank]);
 
     // ---------------------------------------------------------------------//
 

### 2018-10-16 src/main.cpp
Add output file

@@ -384,11 +384,8 @@ int main(int argc, char **argv) {
             }
         }
 
-        scatter_bodies(bodies, sbodies, rank, send_counts, displacements);
-
-        #pragma omp parallel for shared(sbodies)
-        for (size_t i = 0; i < sbodies.size(); i++) {
-            auto& body = sbodies[i];
+        for (size_t i = 0; i < bodies.size(); i++) {
+            auto& body = bodies[i];
 
             if (step % 2 == LEAP) {
                 body.leap(timestep);
@@ -397,8 +394,6 @@ int main(int argc, char **argv) {
                 body.frog(timestep);
             }
         }
-
-        gather_bodies(bodies, sbodies, rank, send_counts, displacements);
 
         t += halfstep;
         step++;

### 2018-10-18 batch.py
Update error logging

@@ -30,7 +30,7 @@ echo '----------------'
 """
 
 # numTimeSteps outputInterval deltaT inputFile
-INVOCATION = "time mpirun -n ${{SLURM_NPROCS}} ./nbody {numTimeSteps} {outputInterval} {deltaT} {inputFile}"
+INVOCATION = "time mpirun -n ${{SLURM_NPROCS}} ./nbody {numTimeSteps} {outputInterval} {deltaT} {inputFile} {enableBarnesHut}"
 
 num_time_steps = 10
 time = 10
@@ -94,7 +94,8 @@ def make_batch(
         numTimeSteps=num_time_steps,
         outputInterval=output_interval,
         deltaT=delta_t,
-        inputFile=input_file
+        inputFile=input_file,
+        enableBarnesHut=(1 if enable_barnes_hut else 0)
     )
 
     return name, batch_args + LOGGING + invocation

### 2018-10-18 src/main.cpp

@@ -19,7 +19,6 @@ MPI_Datatype MPI_Body;
 const unsigned int LEAP = 0;
 const unsigned int FROG = 1;
 
-const bool ENABLE_BARNES_HUT = false;
 const bool ENABLE_LEAPFROG = true;
 const bool ENABLE_LOGGING = false;
 
@@ -241,8 +240,8 @@ std::vector<Body> parse_input_file(std::ifstream& input_fh) {
 }
 
 int main(int argc, char **argv) {
-    if (argc != 5) {
-        fprintf(stdout, "numTimeSteps outputInterval deltaT inputFile\n");
+    if (argc != 6) {
+        fprintf(stdout, "numTimeSteps outputInterval deltaT inputFile enableBarnesHut\n");
         exit(1);
     }
 
@@ -270,6 +269,8 @@ int main(int argc, char **argv) {
 
     const std::string &input_filename = argv[4];
 
+    const bool ENABLE_BARNES_HUT = std::stod(argv[5]) != 0; // big thonk
+
     std::ifstream input_fh(input_filename);
 
     std::vector<Body> bodies = parse_input_file(input_fh);
@@ -293,19 +294,19 @@ int main(int argc, char **argv) {
     displacements.push_back(displacement_acc);
     send_counts.push_back(remainder);
 
-    fprintf(stderr, "Displacements\n");
-    for (size_t i = 0; i < displacements.size(); i++) {
-        fprintf(stderr, "%d ", displacements[i]);
-
-    }
-    fprintf(stderr, "\n");
-
-    fprintf(stderr, "Send counts\n");
-    for (size_t i = 0; i < send_counts.size(); i++) {
-        fprintf(stderr, "%d ", send_counts[i]);
-
-    }
-    fprintf(stderr, "\n");
+    // fprintf(stderr, "Displacements\n");
+    // for (size_t i = 0; i < displacements.size(); i++) {
+    //     fprintf(stderr, "%d ", displacements[i]);
+    //
+    // }
+    // fprintf(stderr, "\n");
+    //
+    // fprintf(stderr, "Send counts\n");
+    // for (size_t i = 0; i < send_counts.size(); i++) {
+    //     fprintf(stderr, "%d ", send_counts[i]);
+    //
+    // }
+    // fprintf(stderr, "\n");
 
     std::vector<Body> sbodies(send_counts[rank]);
 
@@ -409,26 +410,22 @@ int main(int argc, char **argv) {
 
[…]

### 2018-10-18 batch.py
Update batch file

@@ -1,7 +1,7 @@
 from __future__ import print_function
 from itertools import product
 
-BODIES = [16, 64, 256, 1024, 4096]
+BODIES = [4, 16, 64, 256, 1024, 4096]
 NODES = [1, 4, 8, 12]
 TASKS = [1, 4, 8, 12]
 CPUS_PER_TASK = [1, 8, 24]
@@ -9,13 +9,13 @@ ENABLE_BARNES_HUT = [True, False]
 
 BATCH_ARGS = """#!/bin/bash
 #SBATCH --partition=cosc
-#SBATCH --job-name=mbody
+#SBATCH --job-name={name}
 #SBATCH --nodes={nodes}
 #SBATCH --ntasks={ntasks}
 #SBATCH --cpus-per-task={cpus_per_task}
-#SBATCH --error=batcherr/{name}.log
-#SBATCH --output=batchout/{name}.out
-#SBATCH --time=0:10:00
+#SBATCH --error=./batcherr/{name}.log
+#SBATCH --output=./batchout/{name}.out
+#SBATCH --time=0:05:00
 """
 
 LOGGING = """DATE=$(date +"%Y%m%d%H%M")
@@ -114,7 +114,7 @@ def main():
             enable_barnes_hut=enable_barnes_hut
         )
 
-        with open("batches/" + name, "w+") as f:
+        with open("batches/{name}.sh".format(name=name), "w+") as f:
             f.write(batch)
 
 if __name__ == "__main__":

### 2018-10-18 src/main.cpp

@@ -246,7 +246,6 @@ int main(int argc, char **argv) {
     }
 
     // ---------------------------------------------------------------------//
-
     int size;
     int rank;
     MPI_Comm comm = MPI_COMM_WORLD;
@@ -319,9 +318,6 @@ int main(int argc, char **argv) {
         dump_meta_info(num_time_steps, output_interval, timestep, bodies);
         dump_masses(bodies);
         dump_timestep(t, bodies);
-
-        fprintf(stderr, "Barnes-Hut enabled: %s\n", ENABLE_BARNES_HUT ? "true" : "false");
-        fprintf(stderr, "OMP Max threads: %d\n", omp_get_max_threads());
     }
 
     // ---------------------------------------------------------------------//

### 2018-10-18 batch.py
Update README

@@ -15,7 +15,7 @@ BATCH_ARGS = """#!/bin/bash
 #SBATCH --cpus-per-task={cpus_per_task}
 #SBATCH --error=./batcherr/{name}.log
 #SBATCH --output=./batchout/{name}.out
-#SBATCH --time=0:05:00
+#SBATCH --time=0:00:10
 """
 
 LOGGING = """DATE=$(date +"%Y%m%d%H%M")
@@ -30,7 +30,7 @@ echo '----------------'
 """
 
 # numTimeSteps outputInterval deltaT inputFile
-INVOCATION = "time mpirun -n ${{SLURM_NPROCS}} ./nbody {numTimeSteps} {outputInterval} {deltaT} {inputFile} {enableBarnesHut}"
+INVOCATION = "time mpirun -n ${{SLURM_NPROCS}} ./nbody {numTimeSteps} {outputInterval} {deltaT} {inputFile} {enableBarnesHut}\n"
 
 num_time_steps = 10
 time = 10

### 2018-10-18 batch.py
Update batch analysis

@@ -1,5 +1,6 @@
 from __future__ import print_function
 from itertools import product
+import json
 
 BODIES = [4, 16, 64, 256, 1024, 4096]
 NODES = [1, 4, 8, 12]
@@ -38,6 +39,11 @@ fps = 1
 
 output_interval = (num_time_steps / time) / fps
 delta_t = time / num_time_steps
+
+def merge_two_dicts(x, y):
+    z = x.copy()
+    z.update(y)
+    return z
 
 def generate_inputs(bodies):
     with open("batchin/{bodies}.in".format(bodies=bodies), "w+") as f:
@@ -115,6 +121,47 @@ def main():
         )
 
         with open("batches/{name}.sh".format(name=name), "w+") as f:
+            f.write(batch)
+
+def analyse():
+    data = {
+        name: name
+    }
+
+    for (bodies, nodes, tasks, cpus_per_task, enable_barnes_hut)\
+    in product(BODIES, NODES, TASKS, CPUS_PER_TASK, ENABLE_BARNES_HUT):
+        name, batch = make_batch(
+            bodies=bodies,
+            nodes=nodes,
+            tasks=tasks,
+            cpus_per_task=cpus_per_task,
+            enable_barnes_hut=enable_barnes_hut
+        )
+
+
+        with open("./batchout/{name}.out".format(name=name), "w+") as o:
+            text = o.read()
+
+            slurm_details, body = text.split("----------------")
+
+            for line in slurm_details.split('\n')
+                k, v = line.split('=')
+
+                data[k] = v
+
+        with open("./batcherr/{name}.log".format(name=name)) as e:
+            text = e.read()
+            info, time = text.split('\n\n')
+            info_dict = json.loads("{" + info + "}")
+
+
+            _, info_dict["real"] = time.split('\n')[0].split()
+            _, info_dict["user"] = time.split('\n')[1].split()
+            _, info_dict["sys"] = time.split('\n')[1].split()
+
+            data = merge_two_dicts(data, info_dict)
+
+        with open("data.json", "w+") as f:
             f.write(batch)
 
 if __name__ == "__main__":

### 2018-10-18 src/main.cpp

@@ -414,7 +414,7 @@ int main(int argc, char **argv) {
         fprintf(stderr, "numTimeSteps: %d,\n", num_time_steps);
         fprintf(stderr, "outputInterval: %d,\n", output_interval);
         fprintf(stderr, "deltaT: %lf,\n", timestep);
-        fprintf(stderr, "inputFile: %s,\n", input_filename.c_str());
+        fprintf(stderr, "inputFile: \"%s\",\n", input_filename.c_str());
 
         fprintf(stderr, "enableBarnesHut: %d,\n", ENABLE_BARNES_HUT);
 

### 2018-10-18 batch.py
Make a data analysis script

@@ -124,9 +124,7 @@ def main():
             f.write(batch)
 
 def analyse():
-    data = {
-        name: name
-    }
+    data = []
 
     for (bodies, nodes, tasks, cpus_per_task, enable_barnes_hut)\
     in product(BODIES, NODES, TASKS, CPUS_PER_TASK, ENABLE_BARNES_HUT):
@@ -138,6 +136,14 @@ def analyse():
             enable_barnes_hut=enable_barnes_hut
         )
 
+        entry = {
+            name: name
+            bodies: bodies,
+            nodes: nodes,
+            tasks: tasks,
+            cpusPerTask: cpus_per_task,
+            enable_barnes_hut: enable_barnes_hut
+        }
 
         with open("./batchout/{name}.out".format(name=name), "w+") as o:
             text = o.read()
@@ -146,23 +152,22 @@ def analyse():
 
             for line in slurm_details.split('\n')
                 k, v = line.split('=')
-
-                data[k] = v
+                entry[k] = v
 
         with open("./batcherr/{name}.log".format(name=name)) as e:
             text = e.read()
             info, time = text.split('\n\n')
+
+            for line in time.split('\n'):
+                k, v = split()
+                entry[k] = v
+
             info_dict = json.loads("{" + info + "}")
 
+            data.append(merge_two_dicts(entry, info_dict))
 
-            _, info_dict["real"] = time.split('\n')[0].split()
-            _, info_dict["user"] = time.split('\n')[1].split()
-            _, info_dict["sys"] = time.split('\n')[1].split()
-
-            data = merge_two_dicts(data, info_dict)
-
-        with open("data.json", "w+") as f:
-            f.write(batch)
+    with open("data.json", "w+") as f:
+        f.write(json.dumps(data))
 
 if __name__ == "__main__":
     main()

### 2018-10-18 src/main.cpp
Output valid JSON

@@ -406,22 +406,22 @@ int main(int argc, char **argv) {
 
     const double cpu_time_elapsed = cpu_time() - start;
 
-    fprintf(stderr, "%d: %lf,\n", rank, cpu_time_elapsed);
+    fprintf(stderr, "\"%d\": %lf,\n", rank, cpu_time_elapsed);
 
     // ---------------------------------------------------------------------//
 
     if (rank == 0) {
-        fprintf(stderr, "numTimeSteps: %d,\n", num_time_steps);
-        fprintf(stderr, "outputInterval: %d,\n", output_interval);
-        fprintf(stderr, "deltaT: %lf,\n", timestep);
-        fprintf(stderr, "inputFile: \"%s\",\n", input_filename.c_str());
+        fprintf(stderr, "\"numTimeSteps\": %d,\n", num_time_steps);
+        fprintf(stderr, "\"outputInterval\": %d,\n", output_interval);
+        fprintf(stderr, "\"deltaT\": %lf,\n", timestep);
+        fprintf(stderr, "\"inputFile\": \"%s\",\n", input_filename.c_str());
 
-        fprintf(stderr, "enableBarnesHut: %d,\n", ENABLE_BARNES_HUT);
+        fprintf(stderr, "\"enableBarnesHut\": %d,\n", ENABLE_BARNES_HUT);
 
-        fprintf(stderr, "numBodies: %d,\n", static_cast<int>(bodies.size()));
+        fprintf(stderr, "\"numBodies\": %d,\n", static_cast<int>(bodies.size()));
 
-        fprintf(stderr, "ompMaxThreads: %d,\n", omp_get_max_threads());
-        fprintf(stderr, "mpiCommSize: %d,\n", size);
+        fprintf(stderr, "\"ompMaxThreads\": %d,\n", omp_get_max_threads());
+        fprintf(stderr, "\"mpiCommSize\": %d,\n", size);
     }
     
     return 0;

### 2018-10-18 batch.py
Update batch file for better data analysis

@@ -136,33 +136,39 @@ def analyse():
             enable_barnes_hut=enable_barnes_hut
         )
 
+        print("Parsing {name}".format(name=name))
+
         entry = {
-            name: name
-            bodies: bodies,
-            nodes: nodes,
-            tasks: tasks,
-            cpusPerTask: cpus_per_task,
-            enable_barnes_hut: enable_barnes_hut
+            "name": name,
+            "bodies": bodies,
+            "nodes": nodes,
+            "tasks": tasks,
+            "cpusPerTask": cpus_per_task,
+            "enable_barnes_hut": enable_barnes_hut
         }
 
-        with open("./batchout/{name}.out".format(name=name), "w+") as o:
+        with open("./batchout/{name}.out".format(name=name)) as o:
             text = o.read()
 
-            slurm_details, body = text.split("----------------")
+            slurm_details = text.split("----------------")[0]
 
-            for line in slurm_details.split('\n')
-                k, v = line.split('=')
-                entry[k] = v
+            for line in slurm_details.split('\n'):
+                if line != "":
+                    k, v = line.split('=')
+                    entry[k] = v
+
 
         with open("./batcherr/{name}.log".format(name=name)) as e:
             text = e.read()
             info, time = text.split('\n\n')
 
             for line in time.split('\n'):
-                k, v = split()
-                entry[k] = v
+                if line != "":
+                    k, v = line.split()
+                    entry[k] = v
 
-            info_dict = json.loads("{" + info + "}")
+            # :-1 because there's a trailing comma
+            info_dict = json.loads("{" + info[:-1] + "}")
 
             data.append(merge_two_dicts(entry, info_dict))
 
@@ -170,4 +176,9 @@ def analyse():
         f.write(json.dumps(data))
 
 if __name__ == "__main__":
-    main()
+    answer = raw_input("(b)atch or (a)nalyse? ")
+
+    if answer == "b":
+        main()
+    elif answer == "a":
+        analyse()

### 2018-10-18 reports/ms1_report.tex
Update report

@@
 We decompose each dimension of our \texttt{Body}s into arrays, (e.g.
-the $x$ position, the $x$ veolocity, etc.), \texttt{MPI\_Scatterv} these arrays, recompose them into \texttt{Body}s, run \texttt{Leapfrog} on this subset on each node, decompose the subset into arrays, \texttt{MPI\_Gatherv} these arrays back to the root node, and recompose them into $\texttt{Bodys}$.
+the $x$ position, the $x$ veolocity, etc.), \texttt{MPI\_Scatterv} these arrays, recompose them into \texttt{Body}s, run \texttt{Leapfrog} on this subset on each node, decompose the subset into arrays, \texttt{MPI\_Gatherv} these arrays back to the root node, and recompose them into \texttt{Body}s.
 
@@
 
-\medskip
+\section*{Testing}
+
+\subsection*{Changes since MS2}
+
+In MS2, we decomposed each dimension of our \texttt{Body}s into arrays, (e.g.
+the $x$ position, the $x$ veolocity, etc.), \texttt{MPI\_Scatterv}ed these arrays, recomposed them into \texttt{Body}s, ran \texttt{Leapfrog} on this subset on each node, decomposed the subset into arrays, \texttt{MPI\_Gatherv}ed these arrays back to the root node, and recomposed them into \texttt{Body}s.
+We were also allocating a fresh new \texttt{Vector} after \texttt{Body} recomposition.
+This was, frankly, insane, and took a program with very low memory usage and pressure in the serial implementation to one that constantly allocated and deallocated memory on every single step.
+
+By declaring \texttt{MPI\_Datatype MPI\_Body} with \texttt{MPI\_Type\_contiguous(8, MPI\_DOUBLE, \&MPI\_Body)}, we're now able to send and receive subsets of \texttt{Body}s with just a single \texttt{MPI\_Scatterv} and \texttt{MPI\_Gatherv} call.
+We also \texttt{MPI\_Scatterv} and \texttt{MPI\_Gatherv} directly into \texttt{Vector}s allocated at the start of the program, without intermediate array allocations.
+This eliminated transient out-of-memory errors, and likely reduced MPI overhead.
+
+Report content
+   This content adds to the content from previous reports.
+The ECP describes the previous reports as chapters, you can use this opportunity to do a cleanup edit of your previous content but you aren't required to do a major rewrite.
+The content you will be adding should contain:
+    The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
+    Analyse your results. Does your implementation scale? 
+    Summarise your findings for the semester (eg what worked and what didn't) including any limitations you expect scaling up your implementation to solving truely massive problem instances.
+
+Marking criteria: /20
+  Code compiles out of repo /2
+  Testing description       /8
+  Analysis                  /6
+  Summarise findings        /4
 

### 2018-10-18 reports/ms1_report.tex

@@
 
-Report content
-   This content adds to the content from previous reports.
-The ECP describes the previous reports as chapters, you can use this opportunity to do a cleanup edit of your previous content but you aren't required to do a major rewrite.
-The content you will be adding should contain:
-    The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
+\subsection*{Testing Performed}
+
+\begin{enumerate}
+    \item The number of bodies ($n$) (\texttt{4, 16, 64, 256, 1024, 4096})
+    \item \texttt{--nodes} (\texttt{1, 4, 8, 12})
+    \item \texttt{--ntasks} (\texttt{1, 4, 8, 12})
+    \item \texttt{--cpus-per-task} (\texttt{1, 4, 16})
+    \item \texttt{ENABLE\_BARNES\_HUT} (\texttt{true, false})
+\end{enumerate}
+
+All valid combinations\footnote{As an example, \texttt{TASKS = 1, NODES=12} is not an admissable configuration} of the values in the lists above were performed.
+Tweaks in the number of bodies (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), the number of nodes (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI), the number of tasks (it made sense allocate the same number of tasks and nodes)
+
+% The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
     Analyse your results. Does your implementation scale? 

### 2018-10-18 reports/ms1_report.tex
Update report

@@
 
-All valid combinations\footnote{As an example, \texttt{TASKS = 1, NODES=12} is not an admissable configuration} of the values in the lists above were performed.
-Tweaks in the number of bodies (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), the number of nodes (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI), the number of tasks (it made sense allocate the same number of tasks and nodes)
+% The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
+All valid combinations\footnote{As an example, \texttt{TASKS = 1, NODES=12} is not an admissable configuration, as the number of tasks must be greater than the number of nodes} of the values in the lists above were performed.
+Tweaks in the number of bodies ($n$) (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), \texttt{--nodes} (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI) and \texttt{--ntasks} (it made sense allocate the same number of tasks and nodes) were made.
+\texttt{--ntasks-per-node} was removed as a tested variable, as it is dynamically set by Slurm.
+The OMP scheduling policy was unable to be set by program argument, and would have require three seperate binaries compiled with different options - thus, this was also removed as a tested variable.
 
-% The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
-    Analyse your results. Does your implementation scale? 
-    Summarise your findings for the semester (eg what worked and what didn't) including any limitations you expect scaling up your implementation to solving truely massive problem instances.
+% Analyse your results. Does your implementation scale? 
+\subsection*{Scaling}
+
+% Summarise your findings for the semester (eg what worked and what didn't) including any limitations you expect scaling up your implementation to solving truely massive problem instances.
+\subsection*{Summary of Findings}
 

### 2018-10-18 batch.py
Update data

@@ -112,6 +112,9 @@ def main():
 
     for (bodies, nodes, tasks, cpus_per_task, enable_barnes_hut)\
     in product(BODIES, NODES, TASKS, CPUS_PER_TASK, ENABLE_BARNES_HUT):
+        if tasks < nodes:
+            continue
+
         name, batch = make_batch(
             bodies=bodies,
             nodes=nodes,
@@ -139,8 +142,6 @@ def analyse():
             enable_barnes_hut=enable_barnes_hut
         )
 
-        print("Parsing {name}".format(name=name))
-
         entry = {
             "name": name,
             "bodies": bodies,
@@ -150,30 +151,40 @@ def analyse():
             "enable_barnes_hut": enable_barnes_hut
         }
 
-        with open("./batchout/{name}.out".format(name=name)) as o:
-            text = o.read()
+        try:
+            with open("./batchout/{name}.out".format(name=name)) as o:
+                text = o.read()
 
-            slurm_details = text.split("----------------")[0]
+                slurm_details = text.split("----------------")[0]
 
-            for line in slurm_details.split('\n'):
-                if line != "":
-                    k, v = line.split('=')
-                    entry[k] = v
+                for line in slurm_details.split('\n'):
+                    if line != "":
+                        k, v = line.split('=')
+                        entry[k] = v
+        except:
+            print("Failed to parse ./batchout/{name}.out".format(name=name))
+            pass
 
+        try:
+            with open("./batcherr/{name}.log".format(name=name)) as e:
+                text = e.read()
+                info, time = text.split('\n\n')
 
-        with open("./batcherr/{name}.log".format(name=name)) as e:
-            text = e.read()
-            info, time = text.split('\n\n')
+                for line in time.split('\n'):
+                    if line != "":
+                        k, v = line.split()
+                        entry[k] = v
 
-            for line in time.split('\n'):
-                if line != "":
-                    k, v = line.split()
-                    entry[k] = v
-
-            # :-1 because there's a trailing comma
-            info_dict = json.loads("{" + info[:-1] + "}")
+                # :-1 because there's a trailing comma
+                info_dict = json.loads("{" + info[:-1] + "}")
 
             data.append(merge_two_dicts(entry, info_dict))
+            print("Success with {name}".format(name=name))
+            continue
+        except:
+            print("Failed to parse ./batcherr/{name}.log".format(name=name))
+            data.append(entry)
+            continue
 
     with open("data.json", "w+") as f:
         f.write(json.dumps(data))

### 2018-10-18 analysis.py
Update plot and analysis

@@ -8,7 +8,9 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 
-def main():
+from mpl_toolkits.mplot3d import Axes3D  # noqa: F401 unused import
+
+def ms2():
     if len(sys.argv) not in [1 + 1, 1 + 2]:
         print("input [output]")
         exit(1)
@@ -58,5 +60,35 @@ def main():
         else:
             plt.show()
 
+def ms3():
+    if len(sys.argv) not in [1 + 1]:
+        print("input")
+
+    in_filename = sys.argv[1]
+
+    with open(in_filename, "r") as f:
+        data = json.loads(f.read())
+
+        fig = plt.figure()
+        ax = fig.add_subplot(111, projection='3d')
+
+        # For each set of style and range settings, plot n random points in the box
+        # defined by x in [23, 32], y in [0, 100], z in [zlow, zhigh].
+        # for c, m, zlow, zhigh in [('r', 'o', -50, -25), ('b', '^', -30, -5)]:
+
+        xs = [ i["tasks"] for i in data if i["numBodies"] == 1024 ]
+        ys = [ i["nodes"] for i in data if i ["numBodies"] == 1024 ] 
+        zs = [ i["time"] for i in data if i["numBodies"] == 1024 ]
+
+        ax.scatter(xs, ys, zs, c="r", marker="o")
+
+        ax.set_xlabel('Tasks')
+        ax.set_ylabel('Nodes')
+        ax.set_zlabel('Time')
+
+        plt.show()
+
 if __name__ == "__main__":
-    main()
+    # ms2()
+    ms3()
+    

### 2018-10-18 batch.py

@@ -106,7 +106,7 @@ def make_batch(
 
     return name, batch_args + LOGGING + invocation
 
-def main():
+def batch():
     for n in BODIES:
         generate_inputs(n)
 
@@ -168,7 +168,13 @@ def analyse():
             for line in time.split('\n'):
                 if line != "":
                     k, v = line.split()
-                    entry[k] = v
+
+                    m, s = v[:-1].split('m')
+                    total_seconds = (int(m) * 60) + int(s)
+
+                    entry[k] = total_seconds
+
+            entry["time"] = entry["user"] + entry["sys"]
 
             # :-1 because there's a trailing comma
             info_dict = json.loads("{" + info[:-1] + "}")
@@ -182,6 +188,6 @@ if __name__ == "__main__":
     answer = raw_input("(b)atch or (a)nalyse? ")
 
     if answer == "b":
-        main()
+        batch()
     elif answer == "a":
         analyse()

### 2018-10-18 batch.py
Add tasks per node as an option

@@ -5,7 +5,8 @@ import json
 BODIES = [4, 16, 64, 256, 1024, 4096]
 NODES = [1, 4, 8, 12]
 TASKS = [1, 4, 8, 12]
-CPUS_PER_TASK = [1, 4, 16]
+TASKS_PER_NODE = [1, 4, 8]
+CPUS_PER_TASK = [1, 4, 8]
 ENABLE_BARNES_HUT = [True, False]
 
 BATCH_ARGS = """#!/bin/bash
@@ -13,6 +14,7 @@ BATCH_ARGS = """#!/bin/bash
 #SBATCH --job-name={name}
 #SBATCH --nodes={nodes}
 #SBATCH --ntasks={ntasks}
+#SBATCH --ntasks-per-node={tasks_per_node}
 #SBATCH --cpus-per-task={cpus_per_task}
 #SBATCH --error=./batcherr/{name}.log
 #SBATCH --output=./batchout/{name}.out
@@ -21,12 +23,15 @@ BATCH_ARGS = """#!/bin/bash
 
 LOGGING = """DATE=$(date +"%Y%m%d%H%M")
 export OMP_NUM_THREADS=${SLURM_CPUS_PER_TASK}
+echo "DATE=$DATE"
 echo "SLURM_JOB_NAME=$SLURM_JOB_NAME"
 echo "SLURM_JOB_ID=$SLURM_JOB_ID"
+echo "SLURM_JOB_NUM_NODES=$SLURM_JOB_NUM_NODES"
 echo "SLURM_NODELIST=$SLURM_NODELIST"
-echo "DATE=$DATE"
-echo "OMP_NUM_THREADS=$OMP_NUM_THREADS"
+echo "SLURM_NTASKS=$SLURM_NTASKS"
 echo "SLURM_TASKS_PER_NODE=$SLURM_TASKS_PER_NODE"
+echo "SLURM_CPUS_PER_TASK=$SLURM_CPUS_PER_TASK"
+echo "OMP_NUM_THREADS=$OMP_NUM_THREADS"
 echo '----------------'
 """
 
@@ -74,13 +79,15 @@ def make_batch(
     bodies,
     nodes,
     tasks,
+    tasks_per_node,
     cpus_per_task,
     enable_barnes_hut
 ):
-    name = "mbody-b{bodies}-n{nodes}-t{tasks}-cpt{cpus_per_task}".format(
[…]

### 2018-10-18 batch.py
Remove tasks as a batch parameter

@@ -4,7 +4,6 @@ import json
 
 BODIES = [4, 16, 64, 256, 1024, 4096]
 NODES = [1, 4, 8, 12]
-TASKS = [1, 4, 8, 12]
 TASKS_PER_NODE = [1, 4, 8]
 CPUS_PER_TASK = [1, 4, 8]
 ENABLE_BARNES_HUT = [True, False]
@@ -13,7 +12,6 @@ BATCH_ARGS = """#!/bin/bash
 #SBATCH --partition=cosc
 #SBATCH --job-name={name}
 #SBATCH --nodes={nodes}
-#SBATCH --ntasks={ntasks}
 #SBATCH --ntasks-per-node={tasks_per_node}
 #SBATCH --cpus-per-task={cpus_per_task}
 #SBATCH --error=./batcherr/{name}.log
@@ -78,15 +76,13 @@ def generate_inputs(bodies):
 def make_batch(
     bodies,
     nodes,
-    tasks,
     tasks_per_node,
     cpus_per_task,
     enable_barnes_hut
 ):
-    name = "mbody-b{bodies}-n{nodes}-t{tasks}-tpn{tasks_per_node}-cpt{cpus_per_task}".format(
+    name = "mbody-b{bodies}-n{nodes}-tpn{tasks_per_node}-cpt{cpus_per_task}".format(
         bodies=bodies,
         nodes=nodes,
-        tasks=tasks,
         tasks_per_node=task_per_node,
         cpus_per_task=cpus_per_task
     )
@@ -98,7 +94,6 @@ def make_batch(
 
     batch_args = BATCH_ARGS.format(
         nodes=nodes,
-        ntasks=tasks,
         tasks_per_node=task_per_node,
         cpus_per_task=cpus_per_task,
         name=name
@@ -118,12 +113,9 @@ def batch():
     for n in BODIES:
         generate_inputs(n)
 
-    for (bodies, nodes, tasks, tasks_per_node, cpus_per_task, enable_barnes_hut)\
-    in product(BODIES, NODES, TASKS, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
+    for (bodies, nodes, tasks_per_node, cpus_per_task, enable_barnes_hut)\
+    in product(BODIES, TASKS, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
         if cpus_per_task * tasks_per_node > 24:
-            continue
-
[…]

### 2018-10-18 batch.py
Fixed batch generation

@@ -83,7 +83,7 @@ def make_batch(
     name = "mbody-b{bodies}-n{nodes}-tpn{tasks_per_node}-cpt{cpus_per_task}".format(
         bodies=bodies,
         nodes=nodes,
-        tasks_per_node=task_per_node,
+        tasks_per_node=tasks_per_node,
         cpus_per_task=cpus_per_task
     )
 
@@ -94,7 +94,7 @@ def make_batch(
 
     batch_args = BATCH_ARGS.format(
         nodes=nodes,
-        tasks_per_node=task_per_node,
+        tasks_per_node=tasks_per_node,
         cpus_per_task=cpus_per_task,
         name=name
     )
@@ -114,15 +114,14 @@ def batch():
         generate_inputs(n)
 
     for (bodies, nodes, tasks_per_node, cpus_per_task, enable_barnes_hut)\
-    in product(BODIES, TASKS, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
+    in product(BODIES, NODES, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
         if cpus_per_task * tasks_per_node > 24:
             continue
 
         name, batch = make_batch(
             bodies=bodies,
             nodes=nodes,
-            tasks=tasks,
-            tasks_per_node=task_per_node,
+            tasks_per_node=tasks_per_node,
             cpus_per_task=cpus_per_task,
             enable_barnes_hut=enable_barnes_hut
         )
@@ -134,7 +133,7 @@ def analyse():
     data = []
 
     for (bodies, nodes, tasks_per_node, cpus_per_task, enable_barnes_hut)\
-    in product(BODIES, TASKS, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
+    in product(BODIES, NODES, TASKS_PER_NODE, CPUS_PER_TASK, ENABLE_BARNES_HUT):
         if cpus_per_task * tasks_per_node > 24:
             continue
 

### 2018-10-18 batch.py
Fix data parsing

@@ -140,7 +140,7 @@ def analyse():
         name, batch = make_batch(
             bodies=bodies,
             nodes=nodes,
-            tasks_per_node=task_per_node,
+            tasks_per_node=tasks_per_node,
             cpus_per_task=cpus_per_task,
             enable_barnes_hut=enable_barnes_hut
         )
@@ -164,8 +164,8 @@ def analyse():
                     if line != "":
                         k, v = line.split('=')
                         entry[k] = v
-        except:
-            print("Failed to parse ./batchout/{name}.out".format(name=name))
+        except Exception as err:
+            print("Failed ./batcherr/{name}.log with {err}".format(name=name, err=err))
             pass
 
         try:
@@ -177,21 +177,21 @@ def analyse():
                     if line != "":
                         k, v = line.split()
                         entry[k] = v
-
                         m, s = v[:-1].split('m')
-                        total_seconds = (int(m) * 60) + int(s)
-
+                        total_seconds = (float(m) * 60) + float(s)
                         entry[k] = total_seconds
-                        entry["time"] = entry["user"] + entry["sys"]
 
-                    # :-1 because there's a trailing comma
-                    info_dict = json.loads("{" + info[:-1] + "}")
+                entry["time"] = entry["user"] + entry["sys"]
+                # :-1 because there's a trailing comma
+
+                info_dict = json.loads("{" + info[:-1] + "}")
 
                 data.append(merge_two_dicts(entry, info_dict))
+
                 print("Success with {name}".format(name=name))
                 continue
-        except:
-            print("Failed to parse ./batcherr/{name}.log".format(name=name))
+        except Exception as err:
+            print("Failed ./batcherr/{name}.log with {err}".format(name=name, err=err))
             data.append(entry)
             continue
 

### 2018-10-18 analysis.py
Update plot

@@ -72,20 +72,39 @@ def ms3():
         fig = plt.figure()
         ax = fig.add_subplot(111, projection='3d')
 
-        # For each set of style and range settings, plot n random points in the box
-        # defined by x in [23, 32], y in [0, 100], z in [zlow, zhigh].
-        # for c, m, zlow, zhigh in [('r', 'o', -50, -25), ('b', '^', -30, -5)]:
+        bodies = 1024
 
-        xs = [ i["tasks"] for i in data if i["numBodies"] == 1024 ]
-        ys = [ i["nodes"] for i in data if i ["numBodies"] == 1024 ] 
-        zs = [ i["time"] for i in data if i["numBodies"] == 1024 ]
+        valid_data = [ i for i in data if "time" in i and i["bodies"] == bodies ]
 
-        ax.scatter(xs, ys, zs, c="r", marker="o")
+        x = {
+            "label": "CPUs per task",
+            "field": "cpusPerTask"
+        }
 
-        ax.set_xlabel('Tasks')
-        ax.set_ylabel('Nodes')
-        ax.set_zlabel('Time')
+        y = {
+            "label": "nodes",
+            "field": "nodes"
+        }
 
+        for c, m, data, label in [
+            ('b', 'o', [ i for i in valid_data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
+            ('r', '^', [ i for i in valid_data if i["enable_barnes_hut"]], "Barnes-Hut")
+        ]:
+            xs = [ i[x["field"]] for i in data ]
+            ys = [ i[y["field"]] for i in data ]
+            zs = [ i["time"] for i in data ]
+            ax.scatter(xs, ys, zs, c=c, marker=m, label=label)
+
+        ax.set_title("{bodies} bodies, 10 timesteps".format(bodies=bodies))
+        ax.set_xlabel('{label} (n)'.format(label=x["label"]))
+        ax.set_ylabel('{label} (n)'.format(label=y["label"]))
+        ax.set_zlabel('user + sys time (s)')
+
+        print()
+
+        ax.legend()
+
+        plt.savefig("reports/images/{bodies}-{xlabel}-{ylabel}".format(bodies=bodies, xlabel=x["field"], ylabel=y["field"]))
         plt.show()
 
 if __name__ == "__main__":

### 2018-10-18 analysis.py
Become cash money

@@ -61,6 +61,23 @@ def ms2():
             plt.show()
 
 def ms3():
+    BODIES = [4, 16, 64, 256, 1024, 4096]
+
+    cpus_per_task = {
+        "label": "CPUs per task",
+        "field": "cpusPerTask"
+    }
+
+    nodes = {
+        "label": "nodes",
+        "field": "nodes"
+    }
+
+    tasks_per_cpu = {
+        "label": "tasks per node",
+        "field": "tasksPerNode"
+    }
+
     if len(sys.argv) not in [1 + 1]:
         print("input")
 
@@ -68,44 +85,39 @@ def ms3():
 
     with open(in_filename, "r") as f:
         data = json.loads(f.read())
+        valid_data = [ i for i in data if "time" in i ]
 
-        fig = plt.figure()
-        ax = fig.add_subplot(111, projection='3d')
+        for bodies in BODIES:
+            for (x, y) in [
+                (cpus_per_task, nodes),
+                (cpus_per_task, tasks_per_cpu),
+                (nodes, tasks_per_cpu)
+            ]:
+                display_figure(bodies=bodies, data=[i for i in valid_data if i["bodies"] == bodies ], x=x, y=y)
 
-        bodies = 1024
 
-        valid_data = [ i for i in data if "time" in i and i["bodies"] == bodies ]
 
-        x = {
-            "label": "CPUs per task",
-            "field": "cpusPerTask"
-        }
+def display_figure(bodies, data, x, y):
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection='3d')
 
-        y = {
-            "label": "nodes",
-            "field": "nodes"
-        }
+    for c, m, d, label in [
+        ('b', 'o', [ i for i in data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
+        ('r', '^', [ i for i in data if i["enable_barnes_hut"]], "Barnes-Hut")
+    ]:
+        xs = [ i[x["field"]] for i in d ]
+        ys = [ i[y["field"]] for i in d ]
+        zs = [ i["time"] for i in d ]
+        ax.scatter(xs, ys, zs, c=c, marker=m, label=label)
 
-        for c, m, data, label in [
-            ('b', 'o', [ i for i in valid_data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
[…]

### 2018-10-19 batch.py
Update batch params

@@ -3,9 +3,9 @@ from itertools import product
 import json
 
 BODIES = [4, 16, 64, 256, 1024, 4096]
-NODES = [1, 4, 8, 12]
-TASKS_PER_NODE = [1, 4, 8]
-CPUS_PER_TASK = [1, 4, 8]
+NODES = [1, 2, 4, 8, 12]
+TASKS_PER_NODE = [1, 2, 4, 8]
+CPUS_PER_TASK = [1, 2, 4, 8]
 ENABLE_BARNES_HUT = [True, False]
 
 BATCH_ARGS = """#!/bin/bash
@@ -16,7 +16,7 @@ BATCH_ARGS = """#!/bin/bash
 #SBATCH --cpus-per-task={cpus_per_task}
 #SBATCH --error=./batcherr/{name}.log
 #SBATCH --output=./batchout/{name}.out
-#SBATCH --time=0:00:10
+#SBATCH --time=0:00:30
 """
 
 LOGGING = """DATE=$(date +"%Y%m%d%H%M")

### 2018-10-19 reports/ms1_report.tex
Update report

@@
     \item The number of bodies ($n$) (\texttt{4, 16, 64, 256, 1024, 4096})
-    \item \texttt{--nodes} (\texttt{1, 4, 8, 12})
-    \item \texttt{--ntasks} (\texttt{1, 4, 8, 12})
-    \item \texttt{--cpus-per-task} (\texttt{1, 4, 16})
+    \item \texttt{--nodes} (\texttt{1, 2, 4, 8, 12})
+    \item \texttt{--ntasks-per-node} (\texttt{1, 2, 4, 8})
+    \item \texttt{--cpus-per-task} (\texttt{1, 2, 4, 8})
     \item \texttt{ENABLE\_BARNES\_HUT} (\texttt{true, false})
@@
 % The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
-All valid combinations\footnote{As an example, \texttt{TASKS = 1, NODES=12} is not an admissable configuration, as the number of tasks must be greater than the number of nodes} of the values in the lists above were performed.
-Tweaks in the number of bodies ($n$) (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), \texttt{--nodes} (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI) and \texttt{--ntasks} (it made sense allocate the same number of tasks and nodes) were made.
-\texttt{--ntasks-per-node} was removed as a tested variable, as it is dynamically set by Slurm.
+All valid combinations\footnote{As an example, \texttt{--ntasks-per-node=8, cpus-per-task=8} is not an admissable configuration, as this would require 64 CPUs on a single node.
+\texttt{dogmatix} only has nodes with 24 or 28 CPUs} of the values in the lists above were performed.
+Tweaks in the number of bodies ($n$) (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), \texttt{--nodes} (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI) and not setting \texttt{--ntasks} parameter (it made sense to allow Slurm to dynamically allocate the required number of tasks based on our \texttt{--ntasks-per-node} and \texttt{--cpus-per-task} parameters) and adding the \texttt{--ntasks-per-node} parameter.
+Further granuality in the \texttt{--cpus-per-task} parameter was also introduced.
+
 The OMP scheduling policy was unable to be set by program argument, and would have require three seperate binaries compiled with different options - thus, this was also removed as a tested variable.
@@
 
+We measure time with a combination of the \texttt{sys} and \texttt{user} times reported by the \texttt{time} core util.
+
+We can see a few trends emerge
+
+\begin{itemize}
+    \item \texttt{--ntasks-per-node=\{tasks\_per\_node\}} causes time to be an order of magnitude higher, across
+
+\end{itemize}
+- Simulations with Barnes-Hut enables have  lrunning times are smaller than their Barnes-Hut disabled counterparts, except on when TODO
+
+\includegraphics[width=2.2cm]{4-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{4-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{4-nodes-tasksPerNode}
+\includegraphics[width=2.2cm]{16-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{16-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{16-nodes-tasksPerNode}
+\includegraphics[width=2.2cm]{64-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{64-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{64-nodes-tasksPerNode}
+\includegraphics[width=2.2cm]{256-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{256-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{256-nodes-tasksPerNode}
+\includegraphics[width=2.2cm]{1024-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{1024-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{1024-nodes-tasksPerNode}
+\includegraphics[width=2.2cm]{4096-cpusPerTask-nodes}
+\includegraphics[width=2.2cm]{4096-cpusPerTask-tasksPerNode}
+\includegraphics[width=2.2cm]{4096-nodes-tasksPerNode}
+
 % Summarise your findings for the semester (eg what worked and what didn't) including any limitations you expect scaling up your implementation to solving truely massive problem instances.
 \subsection*{Summary of Findings}
+
+Efforts can be made to reduce the MPI overhead.
+Currently \texttt{Body}s are sent in their entirety, at every single simulation step.
+We do not have to send m and the precomputed \texttt{Gm} and \texttt{m}, as each node can recover this information from the input file.
+The `eap' step does not need \texttt{Body::Fx} and \texttt{Body::Fy}, and `frog' does not need \texttt{Body::x} and \texttt{Body::y}.
+Thus, we do not need to \texttt{MPI\_Scatterv} and \texttt{MPI\_Gatherv} this information in preparation for each step.
+This effectively halves the amount of data that would be sent by MPI.
+Efforts would need to be made to determine how much of the MPI overhead is due to a hard bandwidth limit on the interconnects between nodes.
+I hypothesise that MPI overhead scales linearly with data transferred, and that the MPI transmission protocol is responsible for the majority of the overhead at low data transfer quantities.
 

### 2018-10-19 analysis.py

@@ -117,7 +117,10 @@ def display_figure(bodies, data, x, y):
 
     ax.legend()
 
-    plt.savefig("reports/images/{bodies}-{xlabel}-{ylabel}".format(bodies=bodies, xlabel=x["field"], ylabel=y["field"]))
+    name = "{bodies}-{xlabel}-{ylabel}".format(bodies=bodies, xlabel=x["field"], ylabel=y["field"])
+
+    print("\includegraphics[width=2.2cm]{" + name + "}")
+    plt.savefig("reports/images/" + name)
 
 if __name__ == "__main__":
     # ms2()

### 2018-10-19 src/Body.cpp

@@ -42,7 +42,7 @@ void Body::leap(double dt) {
     y = y + (vy * dt);
 }
 
-// velocity, mass, and force
+// velocity, and force
 void Body::frog(double dt) {
     const double ax = Fx / m;
     const double ay = Fy / m;

### 2018-10-19 reports/ms1_report.tex
Better data analysis

@@
 
+We test with 10 \texttt{numTimeSteps}, to drown out the time dedicated to input parsing and upfront memory allocation (which is in the vicinity of 100ms).
+
 % The testing actually performed [Do not rewrite your previous section to match this]. Comment on any differences between your test plan and what you actually executed. Give your results.
@@
 \texttt{dogmatix} only has nodes with 24 or 28 CPUs} of the values in the lists above were performed.
-Tweaks in the number of bodies ($n$) (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), \texttt{--nodes} (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI) and not setting \texttt{--ntasks} parameter (it made sense to allow Slurm to dynamically allocate the required number of tasks based on our \texttt{--ntasks-per-node} and \texttt{--cpus-per-task} parameters) and adding the \texttt{--ntasks-per-node} parameter.
+Tweaks in the number of bodies ($n$) (a step from 4 to 16 bodies is double the body density; tests with 16384 bodies did not terminate), \texttt{--nodes} (\texttt{dogmatix} only has 12 nodes in the \texttt{cosc} partition; it did not make sense to disable MPI), not setting \texttt{--ntasks} parameter (it made sense to allow Slurm to dynamically allocate the required number of tasks based on our \texttt{--ntasks-per-node} and \texttt{--cpus-per-task} parameters) and adding the \texttt{--ntasks-per-node} parameter were made since our MS2 plan.
 Further granuality in the \texttt{--cpus-per-task} parameter was also introduced.
@@
 
-We measure time with a combination of the \texttt{sys} and \texttt{user} times reported by the \texttt{time} core util.
+We measure time with a combination of the \texttt{sys} and \texttt{user} times reported by the \texttt{time} coreutil.
 
-We can see a few trends emerge
+We can see a few trends emerge:
 
-\begin{itemize}
-    \item \texttt{--ntasks-per-node=\{tasks\_per\_node\}} causes time to be an order of magnitude higher, across
+\texttt{--ntasks-per-node=8} causes time to be above an order of magnitude higher, across all results.
+This behaviour is most pronounced when testing with 4096 bodies, across a variable number of nodes.
+We can also see the pronounced effect of Barnes-Hut ($O(n \log n)$) at 4096 bodies.
 
-\end{itemize}
-- Simulations with Barnes-Hut enables have  lrunning times are smaller than their Barnes-Hut disabled counterparts, except on when TODO
+\includegraphics[width=14.2cm]{4096-nodes-tasksPerNode}
+
+The behaviour becomes less significant when increasing the number of CPUs per task, resulting in maximum CPU utilization per node.
+I hypothesize that process context switch overhead is causing this behaviour, as having a greater number of tasks allocated to a node than available CPUs causes the CPU (or CPUs) to multiprocess tasks.
+
+\includegraphics[width=14.2cm]{4096-cpusPerTask-tasksPerNode}
+
+Furthermore, the behaviour is still pronounced at 4 bodies, and the resulting times at \texttt{--ntasks-per-node=8} do not appear to be a function of the number of bodies.
+
+\includegraphics[width=14.2cm]{4-cpusPerTask-tasksPerNode}
 

### 2018-10-19 analysis.py

@@ -93,13 +93,23 @@ def ms3():
                 (cpus_per_task, tasks_per_cpu),
                 (nodes, tasks_per_cpu)
             ]:
-                display_figure(bodies=bodies, data=[i for i in valid_data if i["bodies"] == bodies ], x=x, y=y)
+                for elide_8_tpn in [True, False]:
+                    display_figure(
+                        bodies=bodies, 
+                        data=[i for i in valid_data if i["bodies"] == bodies ], 
+                        x=x, 
+                        y=y,
+                        elide_8_tpn=elide_8_tpn
+                    )
 
 
 
-def display_figure(bodies, data, x, y):
+def display_figure(bodies, data, x, y, elide_8_tpn):
     fig = plt.figure()
     ax = fig.add_subplot(111, projection='3d')
+
+    if elide_8_tpn:
+        data = [ i for i in data if i["tasksPerNode"] != 8 ]
 
     for c, m, d, label in [
         ('b', 'o', [ i for i in data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
@@ -110,14 +120,22 @@ def display_figure(bodies, data, x, y):
         zs = [ i["time"] for i in d ]
         ax.scatter(xs, ys, zs, c=c, marker=m, label=label)
 
-    ax.set_title("{bodies} bodies".format(bodies=bodies))
+    ax.set_title("{bodies} bodies{elide_8_tpn}".format(
+        bodies=bodies, 
+        elide_8_tpn=(", --ntasks-per-node=8 elided" if elide_8_tpn else "")
+    ))
     ax.set_xlabel('{label} (n)'.format(label=x["label"]))
     ax.set_ylabel('{label} (n)'.format(label=y["label"]))
     ax.set_zlabel('user + sys time (s)')
 
     ax.legend()
 
-    name = "{bodies}-{xlabel}-{ylabel}".format(bodies=bodies, xlabel=x["field"], ylabel=y["field"])
+    name = "{bodies}-{xlabel}-{ylabel}{elide_8_tpn}".format(
+        bodies=bodies, 
+        xlabel=x["field"], 
+        ylabel=y["field"],
+        elide_8_tpn=("-n8tpn" if elide_8_tpn else "")
+    )
 
     print("\includegraphics[width=2.2cm]{" + name + "}")
     plt.savefig("reports/images/" + name)

### 2018-10-19 reports/ms1_report.tex
Add more data

@@
 Strangely, $n = 16$ with \textit{Barnes-Hut} enabled took longer than expected, when compared to $n \in \{ 4, 8, 12 \}$. Further investigation is required.
-\begin{figure}[b]
-\caption{Scaling characteristics}
-\centering
-\label{fig:performance}
-\includegraphics[width=\textwidth]{performance}
-\end{figure}
+
+\begin{center}
+\includegraphics[width=0.75\textwidth]{performance}
+\end{center}
 
@@
 
+Further evidence can be seen for this multiprocesing overhead hypothesis, where decreasing the total number of processes allocated to each CPU decreases time.
+
+\includegraphics[width=14.2cm]{1024-cpusPerTask-tasksPerNode-n8tpn}
+
+We can also see that \texttt{--ntasks-per-node=1} runs with consistently good performance. We'll drill into data where\texttt{--ntasks-per-node=1} and further analyse its scaling characteristics.
+
 \includegraphics[width=2.2cm]{4-cpusPerTask-nodes}

### 2018-10-19 analysis.py

@@ -93,22 +93,27 @@ def ms3():
                 (cpus_per_task, tasks_per_cpu),
                 (nodes, tasks_per_cpu)
             ]:
-                for elide_8_tpn in [True, False]:
+                for fil in ["", "elide_8_tpn", "just_1_tpn"]:
                     display_figure(
                         bodies=bodies, 
                         data=[i for i in valid_data if i["bodies"] == bodies ], 
                         x=x, 
                         y=y,
-                        elide_8_tpn=elide_8_tpn
+                        fil=fil
                     )
 
-
-
-def display_figure(bodies, data, x, y, elide_8_tpn):
+def display_figure(bodies, data, x, y, fil):
     fig = plt.figure()
     ax = fig.add_subplot(111, projection='3d')
 
-    if elide_8_tpn:
+    extra=""
+
+    if fil == "elide_8_tpn":
+        extra = ", --ntasks-per-node=8 elided"
+        data = [ i for i in data if i["tasksPerNode"] != 8 ]
+
+    if fil == "just_1_tpn":
+        extra = ", --ntasks-per-node=1"
         data = [ i for i in data if i["tasksPerNode"] != 8 ]
 
     for c, m, d, label in [
@@ -120,9 +125,9 @@ def display_figure(bodies, data, x, y, elide_8_tpn):
         zs = [ i["time"] for i in d ]
         ax.scatter(xs, ys, zs, c=c, marker=m, label=label)
 
-    ax.set_title("{bodies} bodies{elide_8_tpn}".format(
-        bodies=bodies, 
-        elide_8_tpn=(", --ntasks-per-node=8 elided" if elide_8_tpn else "")
+    ax.set_title("{bodies} bodies{extra}".format(
+        bodies=bodies,
+        extra=extra
     ))
     ax.set_xlabel('{label} (n)'.format(label=x["label"]))
     ax.set_ylabel('{label} (n)'.format(label=y["label"]))
@@ -130,11 +135,11 @@ def display_figure(bodies, data, x, y, elide_8_tpn):
 
     ax.legend()
 
-    name = "{bodies}-{xlabel}-{ylabel}{elide_8_tpn}".format(
+    name = "{bodies}-{xlabel}-{ylabel}{extra}".format(
         bodies=bodies, 
         xlabel=x["field"], 
         ylabel=y["field"],
-        elide_8_tpn=("-n8tpn" if elide_8_tpn else "")
[…]

### 2018-10-19 reports/ms1_report.tex
Update a lot of stuff

@@
 Testing focussed primarily on $n > 5000$,\footnote{In MS1, I failed to test simulations with a large number of $n$, as I was focussed on testing the accuracy of the implementation, and generated input files would cause high numbers of collisions that made accuracy assessment difficult.
-Testing $n=10000$, we find a 10x speedup with by using a serial Barnes-Hut implementation (Avg.
-$1.12$s, as opposed to $9.87$s without Barnes-Hut).
-Barnes-Hut afforded a much simpler parallelization scheme, as each body could calculate its force against the quadtree concurrently, without locks} where force and integration calculation overhead dominated thread creation overhead.
+Testing $n=10000$, we find a 10x speedup with by using a serial \textit{Barnes-Hut} implementation (Avg.
+$1.12$s, as opposed to $9.87$s without \textit{Barnes-Hut}).
+\textit{Barnes-Hut} afforded a much simpler parallelization scheme, as each body could calculate its force against the quadtree concurrently, without locks} where force and integration calculation overhead dominated thread creation overhead.
 Parallel speedup vanished around the $n \approx 500$ mark.
@@
 This behaviour is most pronounced when testing with 4096 bodies, across a variable number of nodes.
-We can also see the pronounced effect of Barnes-Hut ($O(n \log n)$) at 4096 bodies.
+We can also see the pronounced effect of \textit{Barnes-Hut} ($O(n \log n)$) at 4096 bodies.
 
@@
 
-\includegraphics[width=14.2cm]{1024-cpusPerTask-tasksPerNode-n8tpn}
+\includegraphics[width=14.2cm]{1024-cpusPerTask-tasksPerNode-elide_8_tpn}
 
@@
 
-\includegraphics[width=2.2cm]{4-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{4-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{4-nodes-tasksPerNode}
-\includegraphics[width=2.2cm]{16-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{16-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{16-nodes-tasksPerNode}
-\includegraphics[width=2.2cm]{64-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{64-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{64-nodes-tasksPerNode}
-\includegraphics[width=2.2cm]{256-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{256-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{256-nodes-tasksPerNode}
-\includegraphics[width=2.2cm]{1024-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{1024-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{1024-nodes-tasksPerNode}
-\includegraphics[width=2.2cm]{4096-cpusPerTask-nodes}
-\includegraphics[width=2.2cm]{4096-cpusPerTask-tasksPerNode}
-\includegraphics[width=2.2cm]{4096-nodes-tasksPerNode}
+We see a large spike in running time when we move from 4 to 8 nodes when testing with 16 bodies.
+This is the only number of bodies that this behaviour is exhibited.
+I have no explanation for this behaviour.
+
+\includegraphics[width=14.2cm]{16-cpusPerTask-nodes-just_1_tpn}
+
+The spike occurs later for 64 bodies, instead at 12 nodes.
+
+\includegraphics[width=14.2cm]{64-cpusPerTask-nodes-just_1_tpn}
+
+At 256 we see worse performance with 2 nodes and 1 CPU per task than with 4 nodes. This behaviour vanishes by 4096 bodies, and instead we see an adding more nodes increases the performance.
+
+\includegraphics[width=5.2cm]{256-cpusPerTask-nodes-just_1_tpn}\includegraphics[width=5.2cm]{1024-cpusPerTask-nodes-just_1_tpn}\includegraphics[width=5.2cm]{4096-cpusPerTask-nodes-just_1_tpn}
+
+Perhaps the only thing that provides a parallel speedup is dedicating more CPUs to each task.
+
+\includegraphics[width=14.2cm]{scaling}
 
@@
 
+\subsubsection*{What Worked}
+
+\textit{Barnes-Hut} provided an effective reduction of complexity.
+Small micro-optimisations made with the assistance of a profiler provided decent performance improvements.
+\textit{Leapfrog} increased simulation accuracy.
+OpenMP provided a parallel speedup (after efforts were made to eliminate locking).
+
+\subsubsection*{What Didn't Work}
+
+ Apart from a few anomolies, adding more nodes degraded performance. Running multiple tasks on the same CPU further degraded performance.
+
+ \subsubsection*{What Next}
+
 Efforts can be made to reduce the MPI overhead.
@@
 We do not have to send m and the precomputed \texttt{Gm} and \texttt{m}, as each node can recover this information from the input file.
-The `eap' step does not need \texttt{Body::Fx} and \texttt{Body::Fy}, and `frog' does not need \texttt{Body::x} and \texttt{Body::y}.
+The `leap' step does not need \texttt{Body::Fx} and \texttt{Body::Fy}, and `frog' does not need \texttt{Body::x} and \texttt{Body::y}.
 Thus, we do not need to \texttt{MPI\_Scatterv} and \texttt{MPI\_Gatherv} this information in preparation for each step.
@@
 
-Marking criteria: /20
-  Code compiles out of repo /2
-  Testing description       /8
-  Analysis                  /6
-  Summarise findings        /4
+The \textit{Barnes-Hut} algorithm provides an effective means of compartmentalizing computations, eliminating the need to \texttt{MPI\_Gatherv} evetything back to the root node after each step.
+Each Quad of the \texttt{QuadTree} could be assigned to a seperate node; data would only be transferred between nodes when bodies stray over the boundry of a specific Quad.
+This would dramatically increase the complexity of the MPI messaging implementation.
 

### 2018-10-19 analysis.py

@@ -73,7 +73,7 @@ def ms3():
         "field": "nodes"
     }
 
-    tasks_per_cpu = {
+    tasks_per_node = {
         "label": "tasks per node",
         "field": "tasksPerNode"
     }
@@ -87,13 +87,29 @@ def ms3():
         data = json.loads(f.read())
         valid_data = [ i for i in data if "time" in i ]
 
+        display_2d_figure(
+            bodies=4096,
+            data=[i for i in valid_data if i["bodies"] == 4096 and i["nodes"] == 1 ],
+            x=cpus_per_task
+        )
+
         for bodies in BODIES:
-            for (x, y) in [
-                (cpus_per_task, nodes),
-                (cpus_per_task, tasks_per_cpu),
-                (nodes, tasks_per_cpu)
-            ]:
-                for fil in ["", "elide_8_tpn", "just_1_tpn"]:
+            for fil in ["", "elide_8_tpn", "just_1_tpn"]:
+                if fil == "just_1_tpn":
+                    display_figure(
+                        bodies=bodies, 
+                        data=[i for i in valid_data if i["bodies"] == bodies ], 
+                        x=cpus_per_task, 
+                        y=nodes,
+                        fil=fil
+                    )
+                    continue
+
+                for (x, y) in [
+                    (cpus_per_task, nodes),
+                    (cpus_per_task, tasks_per_node),
+                    (nodes, tasks_per_node)
+                ]:
                     display_figure(
                         bodies=bodies, 
                         data=[i for i in valid_data if i["bodies"] == bodies ], 
@@ -101,6 +117,33 @@ def ms3():
                         y=y,
                         fil=fil
                     )
+
+
+
+def display_2d_figure(bodies, data, x):
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+
+    for c, m, d, label in [
+        ('b', 'o', [ i for i in data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
+        ('r', '^', [ i for i in data if i["enable_barnes_hut"]], "Barnes-Hut")
+    ]:
+        xs = [ i[x["field"]] for i in d ]
+        ys = [ i["time"] for i in d ]
+        ax.scatter(xs, ys, c=c, marker=m, label=label)
+
[…]

### 2018-10-19 reports/ms1_report.tex
Append everything

@@
 
+We failed to find the threshold where MPI overhead gave way to a parallel speedup. Future testing plans may require even higher numbers of bodies.
+
+Testing with higher counts of bodies caused transient out-of-memory issues that persisted even after requests for more memory per node were made.
+This dramatically reduced the statistical significance of tests with 16384 bodies.
+
  \subsubsection*{What Next}
@@
 
+\section*{Appendix}
+
+\includegraphics[width=4.4cm]{4-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{4-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{4-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{4-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{4-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{4-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{4-cpusPerTask-nodes-just_1_tpn}
+
+\includegraphics[width=4.4cm]{16-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{16-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{16-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{16-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{16-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{16-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{16-cpusPerTask-nodes-just_1_tpn}
+
+\includegraphics[width=4.4cm]{64-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{64-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{64-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{64-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{64-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{64-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{64-cpusPerTask-nodes-just_1_tpn}
+\includegraphics[width=4.4cm]{256-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{256-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{256-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{256-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{256-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{256-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{256-cpusPerTask-nodes-just_1_tpn}
+
+\includegraphics[width=4.4cm]{1024-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{1024-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{1024-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{1024-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{1024-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{1024-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{1024-cpusPerTask-nodes-just_1_tpn}
+
+\includegraphics[width=4.4cm]{4096-cpusPerTask-nodes}
+\includegraphics[width=4.4cm]{4096-cpusPerTask-tasksPerNode}
+\includegraphics[width=4.4cm]{4096-nodes-tasksPerNode}
+\includegraphics[width=4.4cm]{4096-cpusPerTask-nodes-elide_8_tpn}
+\includegraphics[width=4.4cm]{4096-cpusPerTask-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{4096-nodes-tasksPerNode-elide_8_tpn}
+\includegraphics[width=4.4cm]{4096-cpusPerTask-nodes-just_1_tpn}
 \end{document}

### 2018-10-19 analysis.py

@@ -142,7 +142,7 @@ def display_2d_figure(bodies, data, x):
 
     name = "scaling"
 
-    print("\includegraphics[width=2.2cm]{" + name + "}")
+    print("\includegraphics[width=4.4cm]{" + name + "}")
     plt.savefig("reports/images/" + name)
 
 def display_figure(bodies, data, x, y, fil):
@@ -185,7 +185,7 @@ def display_figure(bodies, data, x, y, fil):
         extra=("-{fil}".format(fil=fil) if fil else "")
     )
 
-    print("\includegraphics[width=2.2cm]{" + name + "}")
+    print("\includegraphics[width=4.4cm]{" + name + "}")
     plt.savefig("reports/images/" + name)
 
 if __name__ == "__main__":

### 2018-10-22 analysis.py
Yeet

@@ -78,6 +78,12 @@ def ms3():
         "field": "tasksPerNode"
     }
 
+    bodies = {
+        "label": "bodies",
+        "field": "bodies"
+    }
+
+
     if len(sys.argv) not in [1 + 1]:
         print("input")
 
@@ -87,10 +93,14 @@ def ms3():
         data = json.loads(f.read())
         valid_data = [ i for i in data if "time" in i ]
 
-        display_2d_figure(
+        cpu_scaling(
             bodies=4096,
             data=[i for i in valid_data if i["bodies"] == 4096 and i["nodes"] == 1 ],
             x=cpus_per_task
+        )
+
+        body_scaling(
+            data=[i for i in valid_data if i["nodes"] == 1 ]
         )
 
         for bodies in BODIES:
@@ -118,9 +128,7 @@ def ms3():
                         fil=fil
                     )
 
-
-
-def display_2d_figure(bodies, data, x):
+def cpu_scaling(bodies, data, x):
     fig = plt.figure()
     ax = fig.add_subplot(111)
 
@@ -141,6 +149,34 @@ def display_2d_figure(bodies, data, x):
     ax.legend()
 
     name = "scaling"
+
+    print("\includegraphics[width=4.4cm]{" + name + "}")
+    plt.savefig("reports/images/" + name)
+
+def body_scaling(data):
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+
+    x = {
+        "label": "bodies",
+        "field": "bodies"
+    }
+
+    for c, m, d, label in [
+        ('b', 'o', [ i for i in data if not i["enable_barnes_hut"]], "No Barnes-Hut"), 
+        ('r', '^', [ i for i in data if i["enable_barnes_hut"]], "Barnes-Hut")
+    ]:
+        xs = [ i[x["field"]] for i in d ]
+        ys = [ i["time"] for i in d ]
+        ax.scatter(xs, ys, c=c, marker=m, label=label)
+
+    ax.set_title("--ntasks-per-node=1, --nodes=1")
+
+    ax.set_xlabel('{label} (n)'.format(label=x["label"]))
+    ax.set_ylabel('user + sys time (s)')
+    ax.legend()
+
+    name = "bodies_scaling"
 
     print("\includegraphics[width=4.4cm]{" + name + "}")
     plt.savefig("reports/images/" + name)

### 2018-10-22 generate_inputs.py

@@ -1,19 +1,0 @@
-from random import random
-
-BODIES = 20
-VELOCITY = 100
-
-print(BODIES)
-for i in range(0, BODIES):
-    print(str(220 - 2 * i) + "e14")
-
-print(0.00, 10)
-
-for i in range(0, BODIES):
-    print(
-        0,
-        0 + i * 5,
-        0 + i * 5,
-        0,
-    )
-
